---
title: "AAT Analysis Part 3 - Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 3 - Data Analysis

Prior to starting this section, need to go through Part 1 (QIIME prep) and Part 2 (R prep)
End product from both parts:
#### Final Phyloseq object (phyloseq.final) with 13734 taxa, 224 samples, and 17 variables 

## Overview of Steps:
#### 1. Absolute and Relative Abundance 
#### 2. Alpha Diversity
#### 3. Beta Diversity
#### 4. Beta Distance


## Prep R environment

Set your working directory, activate relevant packages, and re-upload final phyloseq object

```{r, echo=TRUE}

## change to your working directory
setwd("/Users/kathryndalton/Dropbox/AAT_microbiome/16S")

#Bring up Packages
library("phyloseq")
library("vegan")
library("DESeq2")
library(btools)
library("ggplot2")
library("reshape2")
library("tidyr")
library(plyr)  ## need to upload before dpylr 
library("dplyr")
library("qiime2R")
library("tidyverse")
library(ggpubr)
library("cowplot")
library(gridExtra)
library(grid)

## Final phyloseq 
seqtab<-read.table("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/aat-otutable_final.tsv", sep = "\t", row.names = 1)
tree<-read_qza("rooted-tree-aat.qza")
aat_metadata <-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/AAT_workbook_contact.txt") 
taxtable_final<-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/aat-taxtable_final.tsv")

phylo.final<-phyloseq(
  otu_table(seqtab, taxa_are_rows = T), 
  phy_tree(tree$data), 
  tax_table(as.data.frame(taxtable_final) %>% select(-Confidence.y) %>% column_to_rownames("Feature.ID") %>% as.matrix()),
  sample_data(aat_metadata %>% as.data.frame() %>% column_to_rownames("SampleID"))
)
phylo.final

### change Confidence.y to just Confidence if not doing additional manual blast step

```



## Step 1 - Absolute and Relative Abundance



```{r, echo=TRUE}





```






```{r, echo=TRUE}
plot(pressure)
```


