---
title: "AAT Analysis Part 3 - Data Analysis for Sharing Paper"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 3 - Data Analysis for Microbial Sharing Chapter/Paper

Prior to starting this section, need to go through Part 1 (QIIME prep) and Part 2 (R prep)
End product from both parts:
#### Final Phyloseq object (phyloseq.final) with 13734 taxa, 224 samples, and 17 variables 

## Overview of Steps:
#### 1. Prep your environment and working datasets
#### 2. Absolute and Relative Abundance 
#### 3. Alpha Diversity
#### 4. Beta Diversity
#### 5. Beta Distance

This is just the analysis for the sharing paper, refer to Part 4 for analysis related to the vet paper.


## Step 1 - Prep R environment

Set your working directory, activate relevant packages, and re-upload final phyloseq object

```{r, echo=TRUE}

## change to your working directory
setwd("/Users/kathryndalton/Dropbox/AAT_microbiome/16S")

#Bring up Packages
library("phyloseq")
library("vegan")
library("DESeq2")
library(btools)
library("ggplot2")
library("reshape2")
library("tidyr")
library(plyr)  ## need to upload before dpylr 
library("dplyr")
library("qiime2R")
library("tidyverse")
library(ggpubr)
library("cowplot")
library(gridExtra)
library(grid)
library("cluster")

## Final phyloseq 
seqtab<-read.table("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/aat-otutable_final.tsv", sep = "\t", row.names = 1)
tree<-read_qza("rooted-tree-aat.qza")
aat_metadata <-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/AAT_workbook_contact.txt") 
taxtable_final<-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/aat-taxtable_final.tsv")

phylo.final<-phyloseq(
  otu_table(seqtab, taxa_are_rows = T), 
  phy_tree(tree$data), 
  tax_table(as.data.frame(taxtable_final) %>% select(-Confidence.y) %>% column_to_rownames("Feature.ID") %>% as.matrix()),
  sample_data(aat_metadata %>% as.data.frame() %>% column_to_rownames("SampleID"))
)
phylo.final

### change Confidence.y to just Confidence if not doing additional manual blast step

## Generate new variables
sample_data(phylo.final)$hostsite<-paste(sample_data(phylo.final)$host, sample_data(phylo.final)$sampletype)
sample_data(phylo.final)$interv<-ifelse(sample_data(phylo.final)$intervention==0, "Control", "Intervention")

##Quick EDA on dataset to confirm samples are there
table(sample_data(phylo.final)$host_env, useNA = "always")
table(sample_data(phylo.final)$sampletype)
table(sample_data(phylo.final)$hostsite)
table(sample_data(phylo.final)$intervention, sample_data(phylo.final)$interv, useNA = "always")

## Get working datasets
phylo.kd<-subset_samples(phylo.final, !is.na(host))
phylo.kd
phy.kde<-subset_samples(phylo.final, !is.na(host_env))
phy.kde
dog<-subset_samples(phylo.final, host=="dog")
dog
kid<-subset_samples(phylo.final, host=="human")
kid
env <-subset_samples(phylo.final, sampletype=="Dust")
env

### Synchronized Colors from all Plots
colors_pre.post<-c("#E6C81F", "#A31603")
colors_host<-c("#7AB5F5", "#F86F61", "#025205")
colors_host_de<-c("#F86F61", "#025205")
colors_contact<-c("#40a695", "#a067bd", "black")
colors_visit<-c("#F19C4C", "#455E9E", "black")


```



## Step 1 - Absolute and Relative Abundance

These are the analyzes that went directly towards either a) the dissertation chapters or b) the journal articles. All other sub-analyzes can be found under the rough AAT_Analysis_Phyloseq_RelAbun.R script. This includes other EDA figures, more DESeq tests, and further staph sub-analyses.

End tables/figures: 
* rel abund figure by host
* table with rel abund top genera, by host/site
* rel abund figure in kids pre/post & cont/interv
* absol abun Staph species by host


```{r, echo=TRUE}

## quick EDA top genus 
TopNOTUs<- names(sort(taxa_sums(phy.kde), TRUE)[1:20])
phy.kde20 <- prune_taxa(TopNOTUs, phy.kde)
tax_table(phy.kde20) 
plot_bar(phy.kde20, "host_env", fill="sampletype", facet_grid=~Genus)
plot1<-plot_bar(phy.kde20, "host_env", fill="Genus")
plot1
plot1+geom_bar(stat="identity", position="stack")
plot2<-plot_bar(phy.kde20, "hostsite", fill="Genus")
plot2
plot2+geom_bar(stat="identity", position="stack")
plot2<-plot_bar(phy.kde20, "hostsite", fill="Genus")
plot2
plot2+geom_bar(stat="identity", position="stack")
plot2+geom_bar(stat="identity", position="stack") +
  facet_grid(.~interv)
# same thing but another way to evaluate
subset_samples(phy.kde20, host=="human") %>% plot_bar("Genus", "Abundance")
subset_samples(phy.kde20, host=="dog") %>% plot_bar("Genus", "Abundance")
subset_samples(phy.kde20, host_env=="env") %>% plot_bar("Genus", "Abundance")

## Get Relative Abundance
ps <- tax_glom(phy.kde, "Genus")
ps0 <- transform_sample_counts(ps, function(x) x / sum(x))
sample_data(ps0)$hostsite<-paste(sample_data(ps0)$host_env, sample_data(ps0)$sampletype)
## plot all samples
df.all<- psmelt(ps0)
ggplot(df.all, aes(Sample, Abundance, fill = Genus)) +
  geom_col() +
  guides(fill=FALSE) +
  ggtitle("kids, dogs, env")
df.all.kid<-filter(df.all, host=="human")
ggplot(df.all.kid, aes(Sample, Abundance, fill = Genus)) +
  geom_col() +
  guides(fill=FALSE) +
  ggtitle("just kids")
## relative abundance based on host and site
ps1 <- merge_samples(ps0, "hostsite")
ps2 <- transform_sample_counts(ps1, function(x) x / sum(x))
plot_bar(ps2, fill="Genus")
plot_bar(ps2, fill="Phylum")
df <- psmelt(ps2)
ggplot(df, aes(Sample, Abundance, fill = Genus)) +
  geom_col() +
  guides(fill=FALSE)
top_phyla <- df %>%
  group_by(Sample, Genus) %>%
  summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
head(top_phyla)
top20 <- top_phyla$Genus[1:20]
df0 <- df %>%
  mutate(Genus = fct_other(Genus, top20))
ggplot(df0, aes(Sample, Abundance, fill = Genus)) +
  geom_col()
## clean up Genus name
levels(df0$Genus) <- sub("g__", "", levels(df0$Genus))
head(df0)

### For Table with relative abundance by Host/Site
df.above03<-filter(df, Abundance>=0.03) #46 --- use this one
df.above03x<- df.above03 %>% group_by(Sample, Genus) %>% summarise(
  mean(Abundance))
knitr::kable(df.above03x, format="html")


#############################################################################################################
## DESeq Differential Abundance Tests ##

# This is only a few of the tests conducted on the dataset (see full AAT_Analysis_Phyloseq_RelAbun.R for full EDA)

## on figure - marked as sig:
# by all host/site
# staph, strept in kids, staph and moraxella in dogs N, proph and conch in oral, corynebac in P, coryne in dust
# in kids by p/p, c/i:
# strept in low post control, strept in high pre, staph in high post, stept in low pre, staph in low post intervention


## By Host and Site

#Run actual test
#1) convert phyloseq object to DESeq Dataset with dispersions estimated, using experiment design factor 
phy.kde_hostsite = phyloseq_to_deseq2(phy.kde, ~ hostsite) 
#2) get geometric means
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
alpha = 0.0001
geoMeans = apply(counts(phy.kde_hostsite), 1, gm_mean)
phy.kde_hostsite = estimateSizeFactors(phy.kde_hostsite, geoMeans = geoMeans)
#3) does rest of test, using default framework
phy.kde_hostsite = DESeq(phy.kde_hostsite, test="Wald", fitType="parametric") ## fit versus "local" - log regression
#default multiple-inference correction is Benjamini-Hochberg
phy.kde_hostsite_res = results(phy.kde_hostsite, cooksCutoff = FALSE)
summary(phy.kde_hostsite_res$padj)
##using parametric - significant differences hostsite, min padj = 0, mean 0.52, max 0.98
phy.kde_hostsite.1 = DESeq(phy.kde_hostsite, test="Wald", fitType="local") 
phy.kde_hostsite_res.1 = results(phy.kde_hostsite.1, cooksCutoff = FALSE)
summary(phy.kde_hostsite_res.1$padj) # same results min padj = 0, mean 0.52, max 0.98

# change (MAP): condition treated vs untreated", meaning that the estimates are of log2(treated /untreated) 
## The lfcSE gives the standard error of the log2FoldChange
# stat is the Wald statistic: the log2FoldChange divided by lfcSE

sigtab_phy.kde_hostsite = phy.kde_hostsite_res[which(phy.kde_hostsite_res$padj < alpha), ]
sigtab_phy.kde_hostsite = cbind(as(sigtab_phy.kde_hostsite, "data.frame"),
                                as(tax_table(phy.kde)[rownames(sigtab_phy.kde_hostsite), ], "matrix"))
head(sigtab_phy.kde_hostsite)
dim(sigtab_phy.kde_hostsite) ### 115 different taxa by host and site from all taxa

## Just top Genera
sigtab_phy.kde_hostsite_top<-sigtab_phy.kde_hostsite %>% 
  rownames_to_column(var="OTU") %>%
  subset(OTU %in% TopNOTUs) 
dim(sigtab_phy.kde_hostsite_top) ## 11 of the top genera different
## With mean absolute abundances
df.absol<- psmelt(phy.kde)
df.absol.hs <- df.absol %>% 
  group_by(OTU, hostsite) %>% 
  summarise(mean(Abundance)) %>%
  spread(., "hostsite", "mean(Abundance)")
sigtab_phy.kde_hostsite_top <- merge(sigtab_phy.kde_hostsite_top, df.absol.hs, by="OTU", all.x=TRUE, all.y=FALSE)
knitr::kable(sigtab_phy.kde_hostsite_top, format="html")
## Just Staph
sigtab_phy.kde_hostsite_staph<-filter(sigtab_phy.kde_hostsite, Genus=="g__Staphylococcus")
dim(sigtab_phy.kde_hostsite_staph) ## 7 Staph species diifer - Almost all epi or pseudo + 1 schleif
sigtab_phy.kde_hostsite_staph <- merge(sigtab_phy.kde_hostsite_staph, df.absol.hs, by.x=0, by.y="OTU", all.x=TRUE, all.y=FALSE)
knitr::kable(sigtab_phy.kde_hostsite_staph, format="html")


###########################################################################################################################

## Differential Abundance Testing Just within Kids ##

## Just Kids Stratified by Visit Type and Contact - two bars one pre one post
kid.c = subset_samples(kid, intervention==0)
kid.gen.c <- tax_glom(kid.c, "Genus")
kg1.c <- transform_sample_counts(kid.gen.c, function(x) x / sum(x))
# need to create new variable combo pre/post and contact level
sample_data(kg1.c)$prepost_contact<-paste(sample_data(kg1.c)$pre_post, sample_data(kg1.c)$contactscore_ppl)
kg2.c <- merge_samples(kg1.c, "prepost_contact")
kg3.c <- transform_sample_counts(kg2.c, function(x) x / sum(x))
df.kg.c <- psmelt(kg3.c)
top_phyla.kg.c <- df.kg.c %>%
  group_by(Sample, Genus) %>%
  summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
head(top_phyla.kg.c)
top20.kg.c <- top_phyla.kg.c$Genus[1:20]
df0.kg.c <- df.kg.c %>%
  mutate(Genus = fct_other(Genus, top20.kg.c))
levels(df0.kg.c$Genus) <- sub("g__", "", levels(df0.kg.c$Genus))
ggplot(df0.kg.c, aes(Sample, Abundance, fill = Genus)) +
  geom_col()

kid.i = subset_samples(kid, intervention==1)
kid.gen.i <- tax_glom(kid.i, "Genus")
kg1.i <- transform_sample_counts(kid.gen.i, function(x) x / sum(x))
sample_data(kg1.i)$prepost_contact<-paste(sample_data(kg1.i)$pre_post, sample_data(kg1.i)$contactscore_ppl)
kg2.i <- merge_samples(kg1.i, "prepost_contact")
kg3.i <- transform_sample_counts(kg2.i, function(x) x / sum(x))
df.kg.i <- psmelt(kg3.i)
top_phyla.kg.i <- df.kg.i %>%
  group_by(Sample, Genus) %>%
  summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
head(top_phyla.kg.i)
top20.kg.i <- top_phyla.kg.i$Genus[1:20]
df0.kg.i <- df.kg.i %>%
  mutate(Genus = fct_other(Genus, top20.kg.i))
levels(df0.kg.i$Genus) <- sub("g__", "", levels(df0.kg.i$Genus))
ggplot(df0.kg.i, aes(Sample, Abundance, fill = Genus)) +
  geom_col()


### DESeq for Just Kids

sample_data(kid)$contact.time<-paste(sample_data(kid)$pre_post, sample_data(kid)$contactscore_ppl)
sample_data(kid)$contact.time <- relevel(as.factor(sample_data(kid)$contact.time), ref = "PRE 0")
## ALL Kid samples, difference by pre/post and contact level
kid.ct = phyloseq_to_deseq2(kid, ~ contact.time)
geoMeans = apply(counts(kid.ct), 1, gm_mean)
kid.ct = estimateSizeFactors(kid.ct, geoMeans = geoMeans)
kid.ct = DESeq(kid.ct, test="Wald", fitType="parametric")
kid.ct_res = results(kid.ct, contrast=list( c("Intercept", "contact.time_POST.0_vs_PRE.0",
                                                "contact.time_POST.1_vs_PRE.0","contact.time_PRE.1_vs_PRE.0") )
                 , cooksCutoff = FALSE)
summary(kid.ct_res$padj) ### 
sigtab_kid.ct = kid.ct_res[which(kid.ct_res$padj < alpha), ]
sigtab_kid.ct = cbind(as(sigtab_kid.ct, "data.frame"), as(tax_table(kid)[rownames(sigtab_kid.ct), ], "matrix"))
head(sigtab_kid.ct)
dim(sigtab_kid.ct) ### 4 diff taxa in kids by pre/post & contact
## Stratified by Control:
kid.c = subset_samples(kid, intervention==0)
kid.c.ct = phyloseq_to_deseq2(kid.c, ~ contact.time)
geoMeans = apply(counts(kid.c.ct), 1, gm_mean)
kid.c.ct = estimateSizeFactors(kid.c.ct, geoMeans = geoMeans)
kid.c.ct = DESeq(kid.c.ct, test="Wald", fitType="parametric")
kid.c.ct_res = results(kid.c.ct, contrast=list( c("Intercept", "contact.time_POST.0_vs_PRE.0",
                                                "contact.time_POST.1_vs_PRE.0","contact.time_PRE.1_vs_PRE.0") )
                 , cooksCutoff = FALSE)
summary(kid.c.ct_res$padj) ### 
sigtab_kid.c.ct = kid.c.ct_res[which(kid.c.ct_res$padj < alpha), ]
sigtab_kid.c.ct = cbind(as(sigtab_kid.c.ct, "data.frame"), as(tax_table(kid.c)[rownames(sigtab_kid.c.ct), ], "matrix"))
head(sigtab_kid.c.ct)
dim(sigtab_kid.c.ct) ### 5 diff taxa in kid.c by pre/post & contact
## Stratified by Intervention:
kid.i = subset_samples(kid, intervention==1)
kid.i.ct = phyloseq_to_deseq2(kid.i, ~ contact.time)
geoMeans = apply(counts(kid.i.ct), 1, gm_mean)
kid.i.ct = estimateSizeFactors(kid.i.ct, geoMeans = geoMeans)
kid.i.ct = DESeq(kid.i.ct, test="Wald", fitType="parametric")
kid.i.ct_res = results(kid.i.ct, contrast=list( c("Intercept", "contact.time_POST.0_vs_PRE.0",
                                                "contact.time_POST.1_vs_PRE.0","contact.time_PRE.1_vs_PRE.0") )
                 , cooksCutoff = FALSE)
summary(kid.i.ct_res$padj) ### 
sigtab_kid.i.ct = kid.i.ct_res[which(kid.i.ct_res$padj < alpha), ]
sigtab_kid.i.ct = cbind(as(sigtab_kid.i.ct, "data.frame"), as(tax_table(kid.i)[rownames(sigtab_kid.i.ct), ], "matrix"))
head(sigtab_kid.i.ct)
dim(sigtab_kid.i.ct) ### 2 diff taxa in kid.i by pre/post & contact





#######################################################################################################################

## Final Abundance Plots ##

cb_palette<-c("#fb8072", "#E69F00", "#F0E442", "#009E73",  "#56B4E9", "#0072B2", "#D55E00", "#CC79A7",
"#7fc97f", "#e41a1c", "#fdc086", "#C0C0C0")
##Dissertaion Plot
ra.plot<-ggplot(df0, aes(Sample, Abundance, fill = Genus)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = cb_palette, guide=guide_legend(reverse = TRUE)) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "dog Oral swab", "dog Skin Swab Perineal", 
                            "dog Skin Swab Inguinal", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dog Oral", "Dog Perineal", "Dog Inguinal", "Dust")) +
  ylab("Relative Abundance") +
  ggtitle("A. Relative Abundance by Sample Host and Site") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size=12),
    legend.title = element_text(face="bold"),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size=16, face="bold")) +
  annotate("text", x="human Nasopharyngeal swab", y=0.4, label="***", size=5) +
  annotate("text", x="human Nasopharyngeal swab", y=0.55, label="***", size=5) +
  annotate("text", x="dog Nasopharyngeal swab", y=0.55, label="***", size=5) +
  annotate("text", x="dog Nasopharyngeal swab", y=0.23, label="***", size=5) +
  annotate("text", x="dog Oral swab", y=0.45, label="***", size=5) +
  annotate("text", x="dog Oral swab", y=0.2, label="***", size=5) +
  annotate("text", x="dog Skin Swab Perineal", y=0.25, label="***", size=5) +
  annotate("text", x="env Dust", y=0.15, label="***", size=5)
ra.plot

##Plot for Journal Article (just dog N site)
df0.1<-filter(df0, Sample == "dog Nasopharyngeal swab" | Sample =="human Nasopharyngeal swab" | Sample =="env Dust")
ggplot(df0.1, aes(Sample, Abundance, fill = Genus)) +
  geom_col()
ra.plot.1<-ggplot(df0.1, aes(Sample, Abundance, fill = Genus)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = cb_palette, guide=guide_legend(reverse = TRUE)) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dust")) +
  ylab("Relative Abundance") +
  ggtitle("A. Relative Abundance by Sample Host") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size=12),
    legend.title = element_text(face="bold"),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size=16, face="bold")) +
  annotate("text", x="human Nasopharyngeal swab", y=0.4, label="***", size=5) +
  annotate("text", x="human Nasopharyngeal swab", y=0.55, label="***", size=5) +
  annotate("text", x="dog Nasopharyngeal swab", y=0.55, label="***", size=5) +
  annotate("text", x="dog Nasopharyngeal swab", y=0.23, label="***", size=5) +
  annotate("text", x="env Dust", y=0.15, label="***", size=5)
ra.plot.1

## Just Kids Stratified by Visit Type and Contact - two bars one pre one post

### need to make each kid graph look / match and then paste them together into one graph
## get them to match colors in earlier RA plot (all sites) -> combine into one big graph
## then add DESeq results to each plot
cb_palette1<-c("#E69F00", "#F0E442", "#56B4E9", "#0072B2", "#CC79A7", "#e41a1c", "#fdc086", "#954CCC", "#C0C0C0")
kidra.plot.con<-ggplot(df0.kg.c, aes(Sample, Abundance, fill = Genus)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = cb_palette1, guide=guide_legend(reverse = TRUE)) +
  scale_x_discrete(limits=c("PRE 1", "POST 1", "PRE 0", "POST 0"), 
                   labels=c( "High Pre", "High Post", "Low Pre", "Low Post")) +
  theme_bw() +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size=12), 
        plot.title = element_text(hjust = 0.5, size=13, face="bold")) +
  ylab("Relative Abundance") +
  ggtitle("B. Patients in Control Visits")
kidra.plot.con<-kidra.plot.con + annotate("text", x="POST 0", y=0.70, label="***", size=5, color="blue")
kidra.plot.con

cb_palette2<-c("#fb8072", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#7fc97f", "#009E73", "#004b49",
               "#ffc0cb", "#e41a1c", "#fdc086", "#954CCC", "#C0C0C0")
kidra.plot.int<-ggplot(df0.kg.i, aes(Sample, Abundance, fill = Genus)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = cb_palette2, guide=guide_legend(reverse = TRUE)) +
  scale_x_discrete(limits=c("PRE 0", "POST 0", "PRE 1", "POST 1"), 
                   labels=c( "High Pre", "High Post", "Low Pre", "Low Post")) +
  theme_bw() +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size=12), 
        plot.title = element_text(hjust = 0.5, size=13, face="bold")) +
  ylab("Relative Abundance") +
  ggtitle("C. Patients in Intervention Visits")
kidra.plot.int<- kidra.plot.int + 
  annotate("text", x="POST 0", y=0.7, label="***", size=5, color="blue") +
  annotate("text", x="PRE 0", y=0.75, label="***", size=5, color="darkred") +
  annotate("text", x="PRE 1", y=0.65, label="***", size=5, color="darkred") +
  annotate("text", x="POST 1", y=0.5, label="***", size=5, color="blue")
kidra.plot.int

#### COMBINE ###
## Diss figure
ggdraw() +
  draw_plot(ra.plot, x = 0, y = 0, width = 0.54, height = 1) +
  draw_plot(kidra.plot.con, x = 0.55, y = 0.5, width = 0.45, height = 0.5) +
  draw_plot(kidra.plot.int, x = 0.55, y = 0, width = 0.45, height = 0.5)
## Journal figure
ggdraw() +
  draw_plot(ra.plot.1, x = 0, y = 0, width = 0.54, height = 1) +
  draw_plot(kidra.plot.con, x = 0.55, y = 0.5, width = 0.45, height = 0.5) +
  draw_plot(kidra.plot.int, x = 0.55, y = 0, width = 0.45, height = 0.5)



###########################################################################################################

### Staph Species Evaluation ###

staph_kiddog<-subset_taxa(phy.kde, Genus=="g__Staphylococcus")
plot_bar(staph_kiddog, "sampletype", "Abundance", fill="host_env")
plot_bar(staph_kiddog, "Species", "Abundance", fill="host_env")   #### GREAT IMAGE!!!
plot_bar(staph_kiddog, "Species", "Abundance", fill="sampletype", facet_grid="host_env") #not as good but shows S.pseu mainly N
k<-plot_bar(staph_kiddog, "Species", "Abundance", fill="host_env")
k+geom_bar(aes(color=host_env, fill=host_env), stat="identity", position="stack") #### GREAT IMAGE!!! BETTER
d<-plot_bar(staph_kiddog, "Species", "Abundance", fill="sampletype", facet_grid="host_env")
d+geom_bar(aes(color=sampletype, fill=sampletype), stat="identity", position="stack") ## This ones good too
plot_bar(staph_kiddog, "Species", "Abundance", fill="host_env", facet_grid=interv~pre_post)
staph1<-plot_bar(staph_kiddog, "Species", "Abundance", fill="host_env", facet_grid=interv~pre_post)
staph1+geom_bar(aes(color=host_env, fill=host_env), stat="identity", position="stack")
### shows changes in S. pseudo between control and intervention visit well
# but do we want something that shows S. aureus changes better 
## now subset just species we care about
staph_kiddog.aep<-subset_taxa(staph_kiddog, Species=="s__aureus" | Species=="s__epidermidis" | Species=="s__pseudintermedius")
staph_kiddog.aep
staph2<-plot_bar(staph_kiddog.aep, "Species", "Abundance", fill="host_env", facet_grid=interv~pre_post)
staph2+geom_bar(aes(color=host_env, fill=host_env), stat="identity", position="stack")
## decent 
staph3<-plot_bar(staph_kiddog.aep, "Species", "Abundance", fill="host_env", facet_grid=interv~contactscore_ppl)
staph3+geom_bar(aes(color=host_env, fill=host_env), stat="identity", position="stack")
sample_data(staph_kiddog.aep)$contact<-ifelse(sample_data(staph_kiddog.aep)$contactscore_ppl==0, "Low Contact Patients", 
                                              ifelse(sample_data(staph_kiddog.aep)$contactscore_ppl==1, "High Contact Patients",
                                                     ifelse(sample_data(staph_kiddog.aep)$contactscore_ppl=="d", "Dogs & Env", NA)))
table(sample_data(staph_kiddog.aep)$contact, sample_data(staph_kiddog.aep)$host_env)
staph3<-plot_bar(staph_kiddog.aep, "Species", "Abundance", fill="host_env", facet_grid=interv~contact)
absol.staph.plot<-staph3+geom_bar(aes(fill=host_env), stat="identity", position="stack") +
  ylab("Absolute Abundance") +
  scale_x_discrete(breaks = c("s__aureus", "s__epidermidis", "s__pseudintermedius"), 
                   labels = c("S. aureus", "S. epidermidis", "S. pseudintermedius")) +
  scale_fill_manual(values=colors_host, 
                    guide=guide_legend(title = NULL),
                    limits=c("human", "dog", "env"),
                    labels=c("Patients", "Dogs", "Environment")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    plot.title = element_text(hjust = 0.5, face = "bold", size=14), 
    axis.text.x = element_text(size=9.5), 
    legend.text = element_text(size=11),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
absol.staph.plot

```


## Step 3 - Alpha Diversity 

These are the analyzes that went directly towards either a) the dissertation chapters or b) the journal articles. All other sub-analyzes can be found under the rough AAT_Analysis_Phyloseq_Alpha.R script. This includes other EDA figures, more KW tests + linear regression, and further alpha sub-analyses, including comparing to staph abundances and QIIME rarefied alpha results.

End tables/figures: 
* figure overall alpha div by host/site
* figure change in alpha div in kids
* KW test for differences



```{r, echo=TRUE}

###  Overall Distribution ###

alpha<-estimate_richness(phy.kde, measures=c("Observed", "Shannon"))
head(alpha)
alpha<-rownames_to_column(alpha, var="SampleID")
alpha.data<-merge(aat_metadata, alpha, by="SampleID", all.x=FALSE)
head(alpha.data)
## not a great plot - can do better w/ a data.frame using ggplot
plot_richness(phy.kde, measures=c("Observed", "Shannon"), x="host_env", color ="sampletype")
plot_richness(phy.kde, measures=c("Observed", "Shannon"), x="hostsite", color ="pre_post")
plot_richness(phy.kde, measures=c("Observed", "Shannon"), x="hostsite", color ="interv")
# instead use data.frame
ggplot(alpha.data, aes(x=sampletype, y=Shannon, colour=pre_post)) +
  geom_point() +
  ggtitle("Shannon")
ggplot(alpha.data, aes(x=sampletype, y=Observed, colour=pre_post)) +
  geom_point() +
  ggtitle("Observed")
## create new variables
alpha.data$hostsite<-paste(alpha.data$host_env, alpha.data$sampletype)
alpha.data$interv<-ifelse(alpha.data$intervention==0, "Control", "Intervention")
## Add Faith's Metric
f1<-estimate_pd(phy.kde)
head(f1)
summary(f1$PD)
f1 <- rownames_to_column(f1, var="SampleID")
alpha.data<-merge(alpha.data, f1, by="SampleID") 
alpha.data$PD<-ifelse(alpha.data$host=="human" & alpha.data$PD>55, NA, alpha.data$PD)
# Quick EDA
alpha.data %>% group_by(hostsite) %>% summarise(mean(Shannon), mean(Observed), mean(PD))
alpha.data %>% group_by(hostsite, interv) %>% summarise(mean(Shannon), mean(Observed), mean(PD))
ggplot(alpha.data, aes(x=hostsite, y=Shannon, colour=pre_post)) +
  geom_point() + 
  ggtitle("Shannon") +
  facet_grid(interv~.)
ggplot(alpha.data, aes(x=hostsite, y=Observed, colour=pre_post)) +
  geom_point() + 
  ggtitle("Observed") +
  facet_grid(interv~.)
ggplot(alpha.data, aes(x=hostsite, y=PD, color=pre_post)) +
  geom_point() + 
  ggtitle("Faith's PD") +
  facet_grid(interv~.)
total.alpha.plot<-plot_richness(phy.kde, measures=c("Observed", "Shannon"), x="hostsite", color ="pre_post", shape="interv")
total.alpha.plot +
  geom_boxplot() ## messy

###############################################################################################

## Kruskal Wallis Test for Diff in Alpha Diversity

# Shannon
kruskal.test(Shannon ~ host_env, data = alpha.data) ####
kruskal.test(Shannon ~ sampletype, data = alpha.data) ####
kruskal.test(Shannon ~ hostsite, data = alpha.data) ####
kruskal.test(Shannon ~ pre_post, data = alpha.data)
kruskal.test(Shannon ~ interv, data = alpha.data)
kruskal.test(Shannon ~ contactscore_ppl, data = alpha.data)
alpha.kid<-filter(alpha.data, host_env=="human")
alpha.dog.all<-filter(alpha.data, host_env=="dog")
alpha.dog.n<-filter(alpha.data, hostsite=="dog Nasopharyngeal swab")
alpha.env<-filter(alpha.data, host_env=="env")
kruskal.test(Shannon ~ pre_post, data = alpha.kid)
kruskal.test(Shannon ~ interv, data = alpha.kid)  ## slight diff
kruskal.test(Shannon ~ contactscore_ppl, data = alpha.kid) ## sligh diff
kruskal.test(Shannon ~ pre_post, data = alpha.dog.all)
kruskal.test(Shannon ~ interv, data = alpha.dog.all)
kruskal.test(Shannon ~ pre_post, data = alpha.dog.n)
kruskal.test(Shannon ~ interv, data = alpha.dog.n)
kruskal.test(Shannon ~ pre_post, data = alpha.env)
kruskal.test(Shannon ~ interv, data = alpha.env)

# Observed
kruskal.test(Observed ~ host_env, data = alpha.data) ####
kruskal.test(Observed ~ sampletype, data = alpha.data) ####
kruskal.test(Observed ~ hostsite, data = alpha.data) ####
kruskal.test(Observed ~ pre_post, data = alpha.data)
kruskal.test(Observed ~ interv, data = alpha.data)
kruskal.test(Observed ~ contactscore_ppl, data = alpha.data)
kruskal.test(Observed ~ pre_post, data = alpha.kid)
kruskal.test(Observed ~ interv, data = alpha.kid)
kruskal.test(Observed ~ contactscore_ppl, data = alpha.kid) ## slight diff
kruskal.test(Observed ~ pre_post, data = alpha.dog.all)                 
kruskal.test(Observed ~ interv, data = alpha.dog.all)
kruskal.test(Observed ~ pre_post, data = alpha.dog.n)
kruskal.test(Observed ~ interv, data = alpha.dog.n)
kruskal.test(Observed ~ pre_post, data = alpha.env)
kruskal.test(Observed ~ interv, data = alpha.env)

# Faith
kruskal.test(PD ~ host_env, data = alpha.data) ####
kruskal.test(PD ~ sampletype, data = alpha.data) ####
kruskal.test(PD ~ hostsite, data = alpha.data) ####
kruskal.test(PD ~ pre_post, data = alpha.data)
kruskal.test(PD ~ interv, data = alpha.data)
kruskal.test(PD ~ contactscore_ppl, data = alpha.data)
kruskal.test(PD ~ pre_post, data = alpha.kid)
kruskal.test(PD ~ interv, data = alpha.kid)
kruskal.test(PD ~ contactscore_ppl, data = alpha.kid)
kruskal.test(PD ~ pre_post, data = alpha.dog.all)
kruskal.test(PD ~ interv, data = alpha.dog.all)
kruskal.test(PD ~ pre_post, data = alpha.dog.n)
kruskal.test(PD ~ interv, data = alpha.dog.n)
kruskal.test(PD ~ pre_post, data = alpha.env)
kruskal.test(PD ~ interv, data = alpha.env)


##############################################################################################

## Final Alpha Diversity Plots ##

#Shannon
shannon.plot<-ggplot(alpha.data, aes(y=Shannon, x=hostsite)) +
  geom_boxplot(width=0.35, fill="#eaeaea") +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("A. Shannon Alpha Diversity") +
  ylim(0, 7.5) +
  guides(colour=guide_legend(title=NULL), shape=guide_legend(title=NULL)) +
  annotate("text", x="human Nasopharyngeal swab", y=7, 
           label="Kruskal-Wallis Test\nOverall Difference\nHost and Sites\np<0.001 ", size=3) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "dog Oral swab", "dog Skin Swab Perineal", 
                            "dog Skin Swab Inguinal", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dog Oral", "Dog Perineal", "Dog Inguinal", "Dust")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    legend.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 
## use this for future diss chapter

## Observed
obs.plot<-ggplot(alpha.data, aes(y=Observed, x=hostsite)) +
  geom_boxplot(width=0.35) +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("Observed OTUs Alpha Diversity") +
  annotate("text", x="human Nasopharyngeal swab", y=875, 
           label="Kruskal-Wallis Test \nDifferemce Host and Sites \np<0.001 ", size=2.5) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "dog Oral swab", "dog Skin Swab Inguinal", 
                            "dog Skin Swab Perineal", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dog Oral", "Dog Inguinal", "Dog Perineal", "Dust")) +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 

#Faith's PD
faith.plot<-ggplot(alpha.data, aes(y=PD, x=hostsite)) +
  geom_boxplot(width=0.35, fill="#eaeaea") +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("B. Faith's Phylogenetic Alpha Diversity") +
  ylab("Faith") +
  guides(colour=FALSE, shape=FALSE) +
  annotate("text", x="human Nasopharyngeal swab", y=95, 
           label="Kruskal-Wallis Test\nOverall Difference\nHost and Sites\np<0.001 ", size=3) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "dog Oral swab", "dog Skin Swab Perineal", 
                            "dog Skin Swab Inguinal", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dog Oral", "Dog Perineal", "Dog Inguinal", "Dust")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 


## just Dog N for journal article 
#Shannon
shannon.plot2<-ggplot(alpha.data, aes(y=Shannon, x=hostsite)) +
  geom_boxplot(width=0.35, fill="#eaeaea") +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("B. Shannon Alpha Diversity") +
  ylim(0, 7.5) +
  scale_color_manual(values=colors_pre.post) +
  guides(colour=guide_legend(title=NULL), shape=guide_legend(title=NULL)) +
  annotate("text", x="human Nasopharyngeal swab", y=7, 
           label="Kruskal-Wallis Test\nOverall Difference Hosts \np<0.001 ", size=3) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab",  "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dust")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    legend.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 
## Observed
obs.plot2<-ggplot(alpha.data, aes(y=Observed, x=hostsite)) +
  geom_boxplot(width=0.35, fill="#eaeaea") +
  ylim(0, 500) +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("A. Total Taxa") +
  scale_color_manual(values=colors_pre.post) +
  guides(colour=guide_legend(title=NULL), shape=guide_legend(title=NULL)) +
  annotate("text", x="human Nasopharyngeal swab", y=400, 
           label="Kruskal-Wallis Test \nDifference Hosts \np<0.001 ", size=3) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dust")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    legend.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 
# Faith's
faith.plot2<-ggplot(alpha.data, aes(y=PD, x=hostsite)) +
  geom_boxplot(width=0.35, fill="#eaeaea") +
  geom_point(aes(colour=pre_post, shape=interv), size=3) +
  ggtitle("C. Faith's Phylogenetic Alpha Diversity") +
  ylab("Faith") +
  guides(colour=guide_legend(title=NULL), shape=guide_legend(title=NULL)) +
  scale_color_manual(values=colors_pre.post) +
  annotate("text", x="human Nasopharyngeal swab", y=95, 
           label="Kruskal-Wallis Test\nOverall Difference\nHost and Sites\np<0.001 ", size=3) +
  scale_x_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", "env Dust"), 
                   labels=c( "Kid Nasal", "Dog Nasal", "Dust")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=12), 
    legend.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size=14, face="bold")) 


#############################################################################################################

##  Change in Alpha Levels


## Need to Change Kid and Dog data differently since organized in different ways (for dog and env, need to account for longitudinal data)

##Kids
### generate wide dataset combining pre and post
alpha.kid$contact<-ifelse(alpha.kid$contactscore_ppl==0, "Low", 
                         ifelse(alpha.kid$contactscore_ppl==1, "High", NA))
alpha.kid_pre <- filter(alpha.kid, pre_post=="PRE") %>%
  subset(select = c(subjectid, Shannon, Observed, PD, batch))
alpha.kid_post <- filter(alpha.kid, pre_post=="POST") %>%
  subset(select = c(subjectid, Shannon, Observed, PD, batch))
alpha.kid_key<-subset(alpha.kid, select = c(sampletype, subjectid, host, record_ID, 
                                 visit_number, match_week, sample_site, interv, contact)) %>%
  unique()
alpha.chg.kid<-merge(alpha.kid_key, alpha.kid_pre, by="subjectid") %>%
  merge(., alpha.kid_post, by="subjectid", suffixes = c("_pre", "_post"))
alpha.chg.kid <- alpha.chg.kid %>%
  mutate(shan_chg = Shannon_post - Shannon_pre, 
         obs_chg = Observed_post - Observed_pre, 
         fpd_chg = PD_post - PD_pre)
head(alpha.chg.kid)
hist(alpha.chg.kid$shan_chg)
hist(alpha.chg.kid$obs_chg)
hist(alpha.chg.kid$fpd_chg)
summary(alpha.chg.kid$shan_chg)
summary(alpha.chg.kid$obs_chg)
summary(alpha.chg.kid$fpd_chg)
alpha.chg.kid_summary<-alpha.chg.kid %>% 
  group_by(interv, contact) %>%
  summarize(
    min(shan_chg), mean(shan_chg), median(shan_chg), max(shan_chg), 
    min(obs_chg), mean(obs_chg), median(obs_chg), max(obs_chg),
    min(fpd_chg), mean(fpd_chg), median(fpd_chg), max(fpd_chg))
knitr::kable(alpha.chg.kid_summary, format="html")


## Dogs and Env
alpha.dogenv<-filter(alpha.data, !host_env=="human")
alpha.dogenv_pre <- filter(alpha.dogenv, pre_post=="PRE") %>%
  subset(., select = c(uniqueid, Shannon, Observed, PD, batch))
alpha.dogenv_post <- filter(alpha.dogenv, pre_post=="POST")  %>%
  subset(., select = c(uniqueid, Shannon, Observed, PD,batch))
alpha.dogenv_key<-subset(alpha.dogenv, select = c(sampletype, subjectid, host, host_env, record_ID, 
                             visit_number, match_week, sample_site, uniqueid, interv)) %>%
  unique()
alpha.chg.dogenv<-merge(alpha.dogenv_key, alpha.dogenv_pre, by="uniqueid") %>%
  merge(., alpha.dogenv_post, by="uniqueid", suffixes = c("_pre", "_post"))
alpha.chg.dogenv <- alpha.chg.dogenv %>%
  mutate(shan_chg = Shannon_post - Shannon_pre, 
         obs_chg = Observed_post - Observed_pre, 
         fpd_chg = PD_post - PD_pre)
head(alpha.chg.dogenv)
hist(alpha.chg.dogenv$shan_chg)
hist(alpha.chg.dogenv$obs_chg)
hist(alpha.chg.dogenv$fpd_chg)
summary(alpha.chg.dogenv$shan_chg)
summary(alpha.chg.dogenv$obs_chg)
summary(alpha.chg.dogenv$fpd_chg)
alpha.chg.dogenv_summary<-alpha.chg.dogenv %>% 
  group_by(interv, sample_site) %>%
  summarize(
    min(shan_chg), mean(shan_chg), median(shan_chg), max(shan_chg), 
    min(obs_chg), mean(obs_chg), median(obs_chg), max(obs_chg),
    min(fpd_chg), mean(fpd_chg), median(fpd_chg), max(fpd_chg))
knitr::kable(alpha.chg.dogenv_summary, format="html")


# KW stats by alpha change 
kruskal.test(shan_chg ~ interv, data = alpha.chg.kid)
kruskal.test(shan_chg ~ contact, data = alpha.chg.kid)
kruskal.test(obs_chg ~ interv, data = alpha.chg.kid)
kruskal.test(obs_chg ~ contact, data = alpha.chg.kid) ### sig
kruskal.test(fpd_chg ~ interv, data = alpha.chg.kid)
kruskal.test(fpd_chg ~ contact, data = alpha.chg.kid) ### sig

kruskal.test(shan_chg ~ interv, data = alpha.chg.dogenv)
kruskal.test(obs_chg ~ interv, data = alpha.chg.dogenv)
kruskal.test(fpd_chg ~ interv, data = alpha.chg.dogenv)
kruskal.test(shan_chg ~ sample_site, data = alpha.chg.dogenv)
kruskal.test(obs_chg ~ sample_site, data = alpha.chg.dogenv) ### sig
kruskal.test(fpd_chg ~ sample_site, data = alpha.chg.dogenv)
alpha.chg.dog.n<-filter(alpha.chg.dogenv, sample_site=="N")
alpha.chg.dog.mo<-filter(alpha.chg.dogenv, sample_site=="MO")
alpha.chg.dog.i<-filter(alpha.chg.dogenv, sample_site=="I")
alpha.chg.dog.r<-filter(alpha.chg.dogenv, sample_site=="R")
alpha.chg.env<-filter(alpha.chg.dogenv, sample_site=="Dust")
kruskal.test(shan_chg ~ interv, data = alpha.chg.dog.n)
kruskal.test(obs_chg ~ interv, data = alpha.chg.dog.n)  ### slight sig
kruskal.test(fpd_chg ~ interv, data = alpha.chg.dog.n)
kruskal.test(shan_chg ~ interv, data = alpha.chg.dog.mo)
kruskal.test(obs_chg ~ interv, data = alpha.chg.dog.mo)
kruskal.test(fpd_chg ~ interv, data = alpha.chg.dog.mo)
kruskal.test(shan_chg ~ interv, data = alpha.chg.dog.i)
kruskal.test(obs_chg ~ interv, data = alpha.chg.dog.i)
kruskal.test(fpd_chg ~ interv, data = alpha.chg.dog.i)
kruskal.test(shan_chg ~ interv, data = alpha.chg.dog.r)
kruskal.test(obs_chg ~ interv, data = alpha.chg.dog.r)
kruskal.test(fpd_chg ~ interv, data = alpha.chg.dog.r)
kruskal.test(shan_chg ~ interv, data = alpha.chg.env)
kruskal.test(obs_chg ~ interv, data = alpha.chg.env)
kruskal.test(fpd_chg ~ interv, data = alpha.chg.env) ## slight sig



#############################################################################################################

## Final Change Plots

##Kids
alpha.kid_prepost<-alpha.kid %>% group_by(pre_post) %>% summarize(Shannon=mean(Shannon),
                                                                  Observed=mean(Observed), 
                                                                  PD=mean(PD, na.rm =T))
alpha.kid_prepost_interv<-alpha.kid %>% group_by(pre_post, interv) %>% summarize(Shannon=mean(Shannon),
                                                                                 Observed=mean(Observed),
                                                                                 PD=mean(PD, na.rm =T))
alpha.kid_prepost$interv <- c("Overall Mean")
shan.change.kid<-ggplot(alpha.kid, aes(x=pre_post, y=Shannon, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_interv, size=2) +
  geom_line(data=alpha.kid_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.kid_prepost, size=2) +
  geom_line(data=alpha.kid_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~contact) +
  ylab("Shannon Alpha Diversity") +
  ylim(0,6.25) +
  ggtitle("Kid Alpha Shannon Diversity Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
shan.change.kid

obs.change.kid<-ggplot(alpha.kid, aes(x=pre_post, y=Observed, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_interv, size=2) +
  geom_line(data=alpha.kid_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.kid_prepost, size=2) +
  geom_line(data=alpha.kid_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~contact) +
  ylab("Observed OTUs") +
  ylim(0,250) +
  ggtitle("Kid Alpha Observed OTU Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
obs.change.kid

fpd.change.kid<-ggplot(alpha.kid, aes(x=pre_post, y=PD, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_interv, size=2) +
  geom_line(data=alpha.kid_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.kid_prepost, size=2) +
  geom_line(data=alpha.kid_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~contact) +
  ylab("Faith's Alpha Diversity") +
  ylim(0,55) +
  ggtitle("Kid Alpha Faith's Phylogenetic Diversity Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
fpd.change.kid

##Dogs ENV
alpha.dogenv_prepost<-alpha.dogenv %>% group_by(pre_post, sample_site) %>% summarize(Shannon=mean(Shannon),
                                                                  Observed=mean(Observed), 
                                                                  PD=mean(PD, na.rm =T))
alpha.dogenv_prepost_interv<-alpha.dogenv %>% group_by(pre_post, sample_site, interv) %>% summarize(Shannon=mean(Shannon),
                                                                                 Observed=mean(Observed),
                                                                                 PD=mean(PD, na.rm =T))
alpha.dogenv_prepost$interv <- c("Overall Mean")

shan.change.dogenv<-ggplot(alpha.dogenv, aes(x=pre_post, y=Shannon, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost_interv, size=2) +
  geom_line(data=alpha.dogenv_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.dogenv_prepost, size=2) +
  geom_line(data=alpha.dogenv_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=guide_legend(title = NULL)) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~sample_site) +
  ylab("Shannon Alpha Diversity") +
  ylim(0,6.25) +
  ggtitle("Dog and Dust Shannon Alpha Diversity Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
shan.change.dogenv

obs.change.dogenv<-ggplot(alpha.dogenv, aes(x=pre_post, y=Observed, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost_interv, size=2) +
  geom_line(data=alpha.dogenv_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.dogenv_prepost, size=2) +
  geom_line(data=alpha.dogenv_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~sample_site) +
  ylab("Observed OTUs") +
  ylim(0,750) +
  ggtitle("Dog and Dust Observed OTU Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
obs.change.dogenv

fpd.change.dogenv<-ggplot(alpha.dogenv, aes(x=pre_post, y=PD, colour=interv)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost_interv, size=2) +
  geom_line(data=alpha.dogenv_prepost_interv, aes(group=interv), size=1.5) +
  geom_point(data=alpha.dogenv_prepost, size=2) +
  geom_line(data=alpha.dogenv_prepost, aes(group=interv), size=1.5) +
  scale_colour_manual(values = colors_visit, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~sample_site) +
  ylab("Faith Alpha Diversity") +
  ylim(0,80) +
  ggtitle("Dog and Dust Faith's Phylogenetic Alpha Diversity Changes") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 11), 
    axis.title.y = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=11)
  )
fpd.change.dogenv

#Change plot just dog N, colored host, faceted interv
alpha.dogenv_short<-filter(alpha.dogenv, sample_site=="N" | sample_site=="Dust")
alpha.dogenv_prepost2<-alpha.dogenv_short %>% group_by(pre_post, host_env, interv) %>% summarize(Shannon=mean(Shannon),
                                                                  Observed=mean(Observed), 
                                                                  PD=mean(PD, na.rm =T))
shan.change.dogenv2<-ggplot(alpha.dogenv_short, aes(x=pre_post, y=Shannon, colour=host_env)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost2, size=2) +
  geom_line(data=alpha.dogenv_prepost2, aes(group=host_env), size=1.5) +
  scale_colour_manual(values = colors_host_de, guide=guide_legend(title = NULL), 
                      labels=c("Dog", "Environment")) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Shannon Alpha Diversity") +
  ylim(0,6.25) +
  ggtitle("E. Shannon Diversity") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
shan.change.dogenv2

obs.change.dogenv2<-ggplot(alpha.dogenv_short, aes(x=pre_post, y=Observed, colour=host_env)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost2, size=2) +
  geom_line(data=alpha.dogenv_prepost2, aes(group=host_env), size=1.5) +
  scale_colour_manual(values = colors_host_de, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Total Taxa") +
  ylim(0,450) +
  ggtitle("D. Total Taxa") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
obs.change.dogenv2

fpd.change.dogenv2<-ggplot(alpha.dogenv_short, aes(x=pre_post, y=PD, colour=host_env)) +
  geom_point(size=1) +
  geom_line(aes(group=uniqueid), size=0.3) +
  geom_point(data=alpha.dogenv_prepost2, size=2) +
  geom_line(data=alpha.dogenv_prepost2, aes(group=host_env), size=1.5) +
  scale_colour_manual(values = colors_host_de, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Faith's Alpha Diversity") +
  ylim(0,55) +
  ggtitle("F. Faith's Phylogenetic Diversity") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
fpd.change.dogenv2


##For Diss - Kids divide by cont & inter and group by contact level
alpha.kid_prepost1<-alpha.kid %>% group_by(pre_post, interv) %>% summarize(Shannon=mean(Shannon),
                                                                  Observed=mean(Observed), 
                                                                  PD=mean(PD, na.rm =T))
alpha.kid_prepost_contact<-alpha.kid %>% group_by(pre_post, contact, interv) %>%
  summarize(Shannon=mean(Shannon),Observed=mean(Observed), PD=mean(PD, na.rm =T))
alpha.kid_prepost1$contact <- c("Overall Mean")

shan.change.kid.contact<-ggplot(alpha.kid, aes(x=pre_post, y=Shannon, colour=contact)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_contact, size=2) +
  geom_line(data=alpha.kid_prepost_contact, aes(group=contact), size=1.5) +
  geom_point(data=alpha.kid_prepost1, size=2) +
  geom_line(data=alpha.kid_prepost1, aes(group=contact), size=1.5) +
  scale_colour_manual(values = colors_contact, guide=guide_legend(title = NULL), 
                      labels=c("High Contact", "Low Contact", "Overall Mean")) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Shannon Alpha Diversity") +
  ylim(0, 5.5) +
  ggtitle("B. Shannon Diversity") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
shan.change.kid.contact

obs.change.kid.contact<-ggplot(alpha.kid, aes(x=pre_post, y=Observed, colour=contact)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_contact, size=2) +
  geom_line(data=alpha.kid_prepost_contact, aes(group=contact), size=1.5) +
  geom_point(data=alpha.kid_prepost1, size=2) +
  geom_line(data=alpha.kid_prepost1, aes(group=contact), size=1.5) +
  scale_colour_manual(values = colors_contact, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Total Taxa") +
  ylim(0,250) +
  ggtitle("A. Total Taxa") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
obs.change.kid.contact

fpd.change.kid.contact<-ggplot(alpha.kid, aes(x=pre_post, y=PD, colour=contact)) +
  geom_point(size=1) +
  geom_line(aes(group=subjectid), size=0.3) +
  geom_point(data=alpha.kid_prepost_contact, size=2) +
  geom_line(data=alpha.kid_prepost_contact, aes(group=contact), size=1.5) +
  geom_point(data=alpha.kid_prepost1, size=2) +
  geom_line(data=alpha.kid_prepost1, aes(group=contact), size=1.5) +
  scale_colour_manual(values = colors_contact, guide=FALSE) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  facet_grid(.~interv) +
  ylab("Faith's Alpha Diversity") +
  ylim(0,55) +
  ggtitle("C. Faith's Phylogenetic Diversity")+
  theme_bw() +
  theme(
    axis.title.x = element_blank(), 
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 8), 
    axis.title.y = element_text(size = 9),
    plot.title = element_text(hjust = 0.5, size=11, face="bold"),
    strip.background = element_rect(fill="light blue"),
    strip.text = element_text(size=10)
  )
fpd.change.kid.contact


################################################################################################

## Combined Plots

## Diss - just Shannon and Faith's, combined overall alpha level all host/sites, and kids divided cont and interv visits, colored by high low contact

## For paper - want overall Obs, Shannon + FPD, just kid&dog N and env
## want to keep as sim to original diss, so keep just kids - like colored by contact
alpha.kid.plots<-ggarrange(obs.change.kid.contact, shan.change.kid.contact, fpd.change.kid.contact,
          common.legend = TRUE, legend = "bottom", nrow = 1)
alpha.dog.plots<-ggarrange(obs.change.dogenv2, shan.change.dogenv2, fpd.change.dogenv2,
          common.legend = TRUE, legend = "bottom", nrow = 1)

alpha.pt.text1<-paste("         ", "In", "Patients:", sep=" ")
alpha.pt.text2<-paste("         ", "In", "Dogs and Environment:", sep=" ")

alpha.pt.text1 <- ggparagraph(text = alpha.pt.text1, size = 13, face = "bold", color = "black", lineheight = 0.7)
alpha.pt.text2 <- ggparagraph(text = alpha.pt.text2, size = 13, face = "bold", color = "black", lineheight = 0.7)

##### FINAL PLOT !!!! ####
ggarrange(alpha.pt.text1, alpha.kid.plots, alpha.pt.text2, alpha.dog.plots, 
          nrow=4, ncol=1, heights = c(0.1, 1, 0.1, 1))


```


## Step 4 - Beta Diversity  

These are the analyzes that went directly towards either a) the dissertation chapters or b) the journal articles. All other sub-analyzes can be found under the rough AAT_Analysis_Phyloseq_Beta.R script. This includes other EDA figures, and further beta sub-analyses.
Beta Differences will be calculated and analyzed next section, including PERMANOVA tests. 

End tables/figures: 
* figure PCoA by host/site


```{r, echo=TRUE}

#1) Overall distribution PCoA with clustering test

## PCOA EDA
ord1<-ordinate(phy.kde, method = "MDS", distance = "unifrac")
plot_ordination(phy.kde, ord1, type = "samples", color="host_env", shape="pre_post")
plot_ordination(phy.kde, ord1, type = "samples", color="match_week", shape="host_env")
plot_ordination(phy.kde, ord1, type = "samples", color="host_env", shape="sampletype")
plot_ordination(phy.kde, ord1, type = "samples", color="hostsite", shape="pre_post")
##check batch effect
plot_ordination(phy.kde, ord1, type = "samples", color="batch")
plot_ordination(phy.kde, ord1, type = "samples", color="sample_plate")
## faceted
p2<-plot_ordination(phy.kde, ord1, type = "samples", color="host_env", shape="pre_post")
p2+ facet_grid(.~interv)
p2<-plot_ordination(phy.kde, ord1, type = "samples", color="host_env", shape="sampletype")
p2+ facet_grid(.~pre_post)
p2+ facet_grid(.~interv)
p2<-plot_ordination(phy.kde, ord1, type = "samples", color="hostsite", shape="pre_post")
p2+ facet_grid(.~interv)

ord2<-ordinate(phy.kde, method = "MDS", distance = "wunifrac")
plot_ordination(phy.kde, ord2, type = "samples", color="host_env", shape="pre_post")
plot_ordination(phy.kde, ord2, type = "samples", color="match_week", shape="host_env")
plot_ordination(phy.kde, ord2, type = "samples", color="host_env", shape="sampletype")
plot_ordination(phy.kde, ord2, type = "samples", color="hostsite", shape="pre_post")
##check batch effect
plot_ordination(phy.kde, ord2, type = "samples", color="batch")
plot_ordination(phy.kde, ord2, type = "samples", color="sample_plate")
## faceted
p<-plot_ordination(phy.kde, ord2, type = "samples", color="host_env", shape="pre_post")
p+ facet_grid(.~interv)
p<-plot_ordination(phy.kde, ord2, type = "samples", color="host_env", shape="sampletype")
p+ facet_grid(.~pre_post)
p+ facet_grid(.~interv)
p<-plot_ordination(phy.kde, ord2, type = "samples", color="hostsite", shape="pre_post")
p+ facet_grid(.~interv)

#############################
### clean graphs
## Diss (all samples)
## UUF
uuf1<-plot_ordination(phy.kde, ord1, type = "samples", color="hostsite", shape="pre_post")
uuf2<-uuf1 + facet_grid(.~interv) +
  geom_point(size=2.5) +
  ggtitle("Unweighted UniFrac") +
  scale_color_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", 
                                "dog Oral swab", "dog Skin Swab Inguinal", "dog Skin Swab Perineal", "NA Dust"),
    labels=c("Patient Nasal", "Dog Nasal", "Dog Oral", "Dog Inguinal", "Dog Perineal", "Environment\nDust")) +
  guides(shape=guide_legend(title=NULL), color=guide_legend(title=NULL)) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill="lightblue"), 
    strip.text = element_text(size=12), 
    legend.text = element_text(size=12), 
    plot.title = element_text(size=14, face="bold", hjust = 0.5)
  )
uuf2
##WUF
wuf1<-plot_ordination(phy.kde, ord2, type = "samples", color="hostsite", shape="pre_post")
wuf2<-wuf1 + facet_grid(.~interv) +
  geom_point(size=2.5) +
  ggtitle("Weighted UniFrac") +
  scale_color_discrete(limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", 
                                "dog Oral swab", "dog Skin Swab Inguinal", "dog Skin Swab Perineal", "NA Dust"),
    labels=c("Patient Nasal", "Dog Nasal", "Dog Oral", "Dog Inguinal", "Dog Perineal", "Environment\nDust")) +
  guides(shape=guide_legend(title=NULL), color=guide_legend(title=NULL)) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill="lightblue"), 
    strip.text = element_text(size=12), 
    legend.text = element_text(size=12), 
    plot.title = element_text(size=14, face="bold", hjust = 0.5)
  ) 
wuf2
ggarrange(uuf2, wuf2, nrow=2, ncol=1,
          common.legend = TRUE, legend = "bottom")


## Paper (just Dog N)
phy.kde_short<-subset_samples(phy.kde, hostsite == "dog Nasopharyngeal swab" | hostsite =="human Nasopharyngeal swab" | hostsite=="NA Dust")

## UUF
uuf2_short<-uuf1 + facet_grid(.~interv) +
  geom_point(size=2.5) +
  ggtitle("Unweighted UniFrac") +
  scale_color_manual(values = colors_host,
                       limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", 
                                "NA Dust"),
                       labels=c("Patient Nasal", "Dog Nasal", "Environment Dust")) +
  guides(shape=guide_legend(title=NULL), color=guide_legend(title=NULL)) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill="lightblue"), 
    strip.text = element_text(size=12), 
    legend.text = element_text(size=12), 
    plot.title = element_text(size=14, face="bold", hjust = 0.5)
  )
uuf2_short
##WUF
wuf2_short<-wuf1 + facet_grid(.~interv) +
  geom_point(size=2.5) +
  ggtitle("Weighted UniFrac") +
  scale_color_manual(values = colors_host,
                     limits=c("human Nasopharyngeal swab", "dog Nasopharyngeal swab", 
                               "NA Dust"),
                     labels=c("Patient Nasal", "Dog Nasal", "Environment Dust")) +
  guides(shape=guide_legend(title=NULL), color=guide_legend(title=NULL)) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill="lightblue"), 
    strip.text = element_text(size=12), 
    legend.text = element_text(size=12), 
    plot.title = element_text(size=14, face="bold", hjust = 0.5)
  ) 
wuf2_short
ggarrange(uuf2_short, wuf2_short, nrow=2, ncol=1,
          common.legend = TRUE, legend = "bottom")


####################################################################################################################### 
### Clustering
# From the clusGap documentation: The clusGap function from the cluster package calculates a goodness of clustering measure, 
# called the gap statistic. For each number of clusters k, it compares (W(k)) with E^*[(W(k))] where the latter is defined 
# via bootstrapping, i.e. simulating from a reference distribution.
pam1 <- function(x,k) list(cluster = pam(x,k, cluster.only=TRUE))
x = phyloseq:::scores.pcoa(ord1, display="sites")
cluster_uuf<- clusGap(x[, 1:2], FUN=pam1, K.max = 20, B = 50)
cluster_uuf
# higher gap statistic better number of clusters
## looks like ~16 clusters
## doesn't tell WHAT clusters they are just that there are clusters
x = phyloseq:::scores.pcoa(ord2, display="sites")
cluster_wuf<- clusGap(x[, 1:2], FUN=pam1, K.max = 20, B = 50)
cluster_wuf
## either 1 big cluster or 20 smaller clusters?
## UNSURE IF WORTHWHILE 
```


## Step 4 - Beta Diversity Distance 

To calculate beta distance, can get it with phyloseq (phyloseq::distance, then melt into matrix and filter) but don't trust as much as QIIME. In QIIME, can get in oen command (versus issues with multiple code lines in phyloseq) and results are more expected based on PCOA distance (on both QIIME and phyloseq plots).

So steps are to go back to Unix/QIIME and get beta diversity metrics there. Unlike alpha diversity, will not be affected by rarefaction. Jus using kid and dog nasal samples for this analysis, since they are more comparable (versus other dog sites and env)

These are the analyzes that went directly towards either a) the dissertation chapters or b) the journal articles. All other sub-analyzes can be found under the rough AAT_Analysis_Phyloseq_Beta.R and AAT_Analysis_BetaAdonis_5.22.R scripts. This includes other EDA figures, and further sub-analyses

End tables/figures: 
* figure beta distances by host/site - seperate contact & visit type
* figure beta distances by host/site - faceted contact & visit type
* adonis tests


```{bash, eval=FALSE, echo=TRUE}

## Steps 1 in QIIME (back in bash)
## Prep
source activate qiime2-2019.7
qiime --help
cd /Users/kathryndalton/Dropbox/AAT_microbiome/16S
## QIIME files at end Part 1: aat-table.qza, aat-rep-seqs.qza, aat-gg-taxonomy.qza, rooted-tree-aat.qza, unrooted-tree-aat.qza

## Step 1 - get beta metrics
qiime feature-table filter-samples \
--i-table aat-featuretable-final.qza \
--m-metadata-file AAT_workbook.txt \
--p-where "[sampletype]='Nasopharyngeal swab'" \
--o-filtered-table new-aat-analysis-5.10/aat-table-nasal.qza
qiime diversity beta-phylogenetic \
--i-table new-aat-analysis-5.10/aat-table-nasal.qza \
--i-phylogeny rooted-tree-aat.qza \
--p-metric 'unweighted_unifrac' \
--o-distance-matrix new-aat-analysis-5.10/unweighted_unifrac_kiddog.qza
qiime diversity beta-phylogenetic \
--i-table new-aat-analysis-5.10/aat-table-nose1.qza \
--i-phylogeny rooted-tree-aat.qza \
--p-metric 'weighted_unifrac' \
--o-distance-matrix new-aat-analysis-5.10/weighted_unifrac_kiddog.qza

#Step 2 - get beta significance files
qiime diversity beta-group-significance \
--i-distance-matrix new-aat-analysis-5.10/unweighted_unifrac_kiddog.qza \
--m-metadata-file AAT_workbook.txt \
--m-metadata-column host \
--o-visualization new-aat-analysis-5.10/unweighted_unifrac_kiddog.qzv \
--p-pairwise
qiime diversity beta-group-significance \
--i-distance-matrix new-aat-analysis-5.10/weighted_unifrac_kiddog.qza \
--m-metadata-file AAT_workbook.txt \
--m-metadata-column host \
--o-visualization new-aat-analysis-5.10/weighted_unifrac_kiddog.qzv \
--p-pairwise

## Step 3 - extract files
mkdir beta_significance
qiime tools extract \
  --input-path new-aat-analysis-5.10/unweighted_unifrac_kiddog.qzv \
  --output-path beta_significance
qiime tools extract \
  --input-path new-aat-analysis-5.10/weighted_unifrac_kiddog.qzv \
  --output-path beta_significance
## copy where final extracted fiel went (will show up in confirm message)
## Should be beta_significance/9c016d13-060e-4cd2-bb69-a6aa9539a1f2 (unweighted) and beta_significance/617045d2-c098-457b-8866-8bb0aed949fd (weighted)  --- will say in unix console
## confirm alphanumeric filepath
## move the file we want (raw data with all distances) into main directory
mv beta_significance/9c016d13-060e-4cd2-bb69-a6aa9539a1f2/data/raw_data.tsv unweighted_unifrac_kiddog.tsv
mv beta_significance/617045d2-c098-457b-8866-8bb0aed949fd/data/raw_data.tsv weighted_unifrac_kiddog.tsv

```


### Back in R console

Creating Figures and Running Adonis PERMANOVA test
adonis = PERMANOVA for beta group significance. Can use formula to test multivariate analysis, including interaction between covariables. Used to determine whether groups of samples are significantly different from one another using the ADONIS permutation-based test and test if there is a statistically signficant difference between our sample types. One way to do this is with the betadisper and adonis functions from the vegan package. adonis can tell us if there is a statistical difference between groups, but it has an assumption that must be met that we first need to check with betadisper, and that is that there is a sufficient level of homogeneity of dispersion within groups. If there is not, then adonis can be unreliable.


``` {r, echo=TRUE} 

## Upload distance files
uuf_kiddog<-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/unweighted_unifrac_kiddog.tsv")
wuf_kiddog<-read_tsv("/Users/kathryndalton/Dropbox/AAT_microbiome/16S/weighted_unifrac_kiddog.tsv")

######### UNWEIGHTED #############

## setting up dataframes and summary of distance results ### 
uuf_kiddog$X1<-NULL
merge1 = merge(uuf_kiddog, aat_metadata, by.x = "SubjectID1", by.y="SampleID") #4639 x 21
uuf.distance = merge(merge1, aat_metadata, by.x = "SubjectID2", by.y="SampleID", suffix=c(".1", ".2")) #4639 x 37
uuf.distance$prepostmatch<-ifelse(uuf.distance$pre_post.1=="PRE"&uuf.distance$pre_post.2=="PRE", "PRE",
                                  ifelse(uuf.distance$pre_post.1=="POST"&uuf.distance$pre_post.2=="POST", "POST", "mismatch"))
table(uuf.distance$prepostmatch)
uuf.distance<-filter(uuf.distance, pre_post.1=="PRE"&pre_post.2=="PRE" | pre_post.1=="POST"&pre_post.2=="POST")
table(uuf.distance$prepostmatch)
uuf.distance$visitmatch<-ifelse(uuf.distance$match_week.1==uuf.distance$match_week.2, "Same Visit", "Different Visit")
table(uuf.distance$visitmatch)
table(uuf.distance$visitmatch, uuf.distance$prepostmatch)
uuf.distance$visitprepostmatch<-as.factor(paste(uuf.distance$visitmatch, uuf.distance$prepostmatch))
table(uuf.distance$visitprepostmatch)
## for visit type - don't really care about mismatch, just want it to stratify, will remove for graphing
uuf.distance$visittype<-ifelse(uuf.distance$intervention.1==0&uuf.distance$intervention.2==0, "Control",
                               ifelse(uuf.distance$intervention.1==1&uuf.distance$intervention.2==1, "Intervention", "mismatch"))
table(uuf.distance$visittype, uuf.distance$visitprepostmatch)
## for contact score - for sake of ease, "High" means either one or both kids have high dog contact in kid-kid difference
## -> ran analysis seperating out and found no diiference pre vs post in  both-high or one-high-one-low (both sig)
uuf.kid.full<-filter(uuf.distance, host.1=="human"&host.2=="human") ## just kid-kid comparisons
uuf.kid.full$contact<-ifelse(uuf.kid.full$contactscore_ppl.1==0 & uuf.kid.full$contactscore_ppl.2==0, "Low", "High")
table(uuf.kid.full$contact)
table(uuf.kid.full$contact, uuf.kid.full$visitprepostmatch)
uuf.kiddog.full <- filter(uuf.distance, host.1=="human"&host.2=="dog" | host.1=="dog"&host.2=="human") # just kid-dog comparisons
uuf.kiddog.full$contact<-ifelse(uuf.kiddog.full$contactscore_ppl.1==0 | uuf.kiddog.full$contactscore_ppl.2==0, "Low",
                                ifelse(uuf.kiddog.full$contactscore_ppl.1==1 | uuf.kiddog.full$contactscore_ppl.2==1, "High", "mismatch"))
table(uuf.kiddog.full$contact)
table(uuf.kiddog.full$contact, uuf.kiddog.full$visitprepostmatch)

uuf.kid.full %>% group_by(prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kid.full %>% group_by(visitmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kid.full %>% group_by(visitprepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kid.full %>% group_by(visittype, contact) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kid.full %>% filter(visittype=="Control") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kid.full %>% filter(visittype=="Intervention") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))

uuf.kiddog.full %>% group_by(prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kiddog.full %>% group_by(visitmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kiddog.full %>% group_by(visitprepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kiddog.full %>% group_by(visittype, contact) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kiddog.full %>% filter(visittype=="Control") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
uuf.kiddog.full %>% filter(visittype=="Intervention") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))


####################################################################################################
######## WEIGHTED #############

## setting up dataframes and summary of distance results ### 
wuf_kiddog$X1<-NULL
merge1 = merge(wuf_kiddog, aat_metadata, by.x = "SubjectID1", by.y="SampleID") #4639 x 21
wuf.distance = merge(merge1, aat_metadata, by.x = "SubjectID2", by.y="SampleID", suffix=c(".1", ".2")) #4639 x 37
wuf.distance$prepostmatch<-ifelse(wuf.distance$pre_post.1=="PRE"&wuf.distance$pre_post.2=="PRE", "PRE",
                                  ifelse(wuf.distance$pre_post.1=="POST"&wuf.distance$pre_post.2=="POST", "POST", "mismatch"))
table(wuf.distance$prepostmatch)
wuf.distance<-filter(wuf.distance, pre_post.1=="PRE"&pre_post.2=="PRE" | pre_post.1=="POST"&pre_post.2=="POST")
table(wuf.distance$prepostmatch)
wuf.distance$visitmatch<-ifelse(wuf.distance$match_week.1==wuf.distance$match_week.2, "Same Visit", "Different Visit")
table(wuf.distance$visitmatch)
table(wuf.distance$visitmatch, wuf.distance$prepostmatch)
wuf.distance$visitprepostmatch<-as.factor(paste(wuf.distance$visitmatch, wuf.distance$prepostmatch))
table(wuf.distance$visitprepostmatch)
wuf.distance$visittype<-ifelse(wuf.distance$intervention.1==0&wuf.distance$intervention.2==0, "Control",
                               ifelse(wuf.distance$intervention.1==1&wuf.distance$intervention.2==1, "Intervention", "mismatch"))
table(wuf.distance$visittype, wuf.distance$visitprepostmatch)


wuf.kid.full<-filter(wuf.distance, host.1=="human"&host.2=="human") ## just kid-kid comparisons
wuf.kid.full$contact<-ifelse(wuf.kid.full$contactscore_ppl.1==0 & wuf.kid.full$contactscore_ppl.2==0, "Low", "High")
table(wuf.kid.full$contact)
table(wuf.kid.full$contact, wuf.kid.full$visitprepostmatch)
wuf.kiddog.full <- filter(wuf.distance, host.1=="human"&host.2=="dog" | host.1=="dog"&host.2=="human") # just kid-dog comparisons
wuf.kiddog.full$contact<-ifelse(wuf.kiddog.full$contactscore_ppl.1==0 | wuf.kiddog.full$contactscore_ppl.2==0, "Low",
                                ifelse(wuf.kiddog.full$contactscore_ppl.1==1 | wuf.kiddog.full$contactscore_ppl.2==1, "High", "mismatch"))
table(wuf.kiddog.full$contact, wuf.kiddog.full$visitprepostmatch)

wuf.kid.full %>% group_by(prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kid.full %>% group_by(visitmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kid.full %>% group_by(visitprepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kid.full %>% group_by(visittype, contact) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kid.full %>% filter(visittype=="Control") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kid.full %>% filter(visittype=="Intervention") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))

wuf.kiddog.full %>% group_by(prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kiddog.full %>% group_by(visitmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kiddog.full %>% group_by(visitprepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kiddog.full %>% group_by(visittype, contact) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kiddog.full %>% filter(visittype=="Control") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))
wuf.kiddog.full %>% filter(visittype=="Intervention") %>% group_by(contact, prepostmatch) %>% 
  summarize(dist_mean=mean(Distance), dist_median=median(Distance), min=min(Distance), max=max(Distance), iqr=IQR(Distance))



############################################################################################################

#### Adonis ###

### UUF #####
#  Get one-way Adonis results each analysis - kid-kid
adonis(Distance ~ prepostmatch, data = uuf.kid.full, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kid.full, perm=9999)  
adonis(Distance ~ contact, data = uuf.kid.full, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = uuf.kid.full, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kid.full, perm=9999)  
#stratified - 
## by visittype
uuf.kid.c<-filter(uuf.kid.full, visittype=="Control") 
adonis(Distance ~ prepostmatch, data = uuf.kid.c, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kid.c, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = uuf.kid.c, perm=9999) 
adonis(Distance ~ contact, data = uuf.kid.c, perm=9999) 
uuf.kid.i<-filter(uuf.kid.full, visittype=="Intervention") 
adonis(Distance ~ prepostmatch, data = uuf.kid.i, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kid.i, perm=9999)
adonis(Distance ~ visitprepostmatch, data = uuf.kid.i, perm=9999) 
adonis(Distance ~ contact, data = uuf.kid.i, perm=9999) 
## don't care about mismatched visits eg. kid Cont - kid interv
## by contact
uuf.kid.h<-filter(uuf.kid.full, contact=="High") 
adonis(Distance ~ prepostmatch, data = uuf.kid.h, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kid.h, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = uuf.kid.h, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kid.h, perm=9999) 
uuf.kid.l<-filter(uuf.kid.full, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kid.l, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kid.l, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = uuf.kid.l, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kid.l, perm=9999) 

####  Get one-way adonis results each analysis - kid-dog
adonis(Distance ~ prepostmatch, data = uuf.kiddog.full, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kiddog.full, perm=9999)  
adonis(Distance ~ contact, data = uuf.kiddog.full, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = uuf.kiddog.full, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kiddog.full, perm=9999)  
#stratified - 
## by visittype
uuf.kiddog.c<-filter(uuf.kiddog.full, visittype=="Control") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.c, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kiddog.c, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = uuf.kiddog.c, perm=9999) 
adonis(Distance ~ contact, data = uuf.kiddog.c, perm=9999) 
uuf.kiddog.i<-filter(uuf.kiddog.full, visittype=="Intervention") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.i, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kiddog.i, perm=9999)
adonis(Distance ~ visitprepostmatch, data = uuf.kiddog.i, perm=9999) 
adonis(Distance ~ contact, data = uuf.kiddog.i, perm=9999) 
## by contact
uuf.kiddog.h<-filter(uuf.kiddog.full, contact=="High") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.h, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kiddog.h, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = uuf.kiddog.h, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kiddog.h, perm=9999) 
uuf.kiddog.l<-filter(uuf.kiddog.full, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.l, perm=9999) 
adonis(Distance ~ visitmatch, data = uuf.kiddog.l, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = uuf.kiddog.l, perm=9999) 
adonis(Distance ~ visittype, data = uuf.kiddog.l, perm=9999) 

### multivariate adonis kid-kid
adonis(Distance ~ prepostmatch + visitmatch, data=uuf.kid.full, perm=9999) 
adonis(Distance ~ prepostmatch * visitmatch, data=uuf.kid.full, perm=9999)
adonis(Distance ~ prepostmatch + visitmatch + contact + visittype, data=uuf.kid.full, perm=9999)
### multivariate adonis kid-dog
adonis(Distance ~ prepostmatch + visitmatch, data=uuf.kiddog.full, perm=9999)
adonis(Distance ~ prepostmatch * visitmatch, data=uuf.kiddog.full, perm=9999)
adonis(Distance ~ prepostmatch + visitmatch + contact + visittype, data=uuf.kiddog.full, perm=9999)

## Double Stratified Adonis 
uuf.kid.ch<-filter(uuf.kid.c, contact=="High") 
uuf.kid.cl<-filter(uuf.kid.c, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kid.ch, perm=9999) #p<0.0001
adonis(Distance ~ prepostmatch, data = uuf.kid.cl, perm=9999) #p=0.6924
uuf.kid.ih<-filter(uuf.kid.i, contact=="High") 
uuf.kid.il<-filter(uuf.kid.i, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kid.ih, perm=9999) #p=0.0003
adonis(Distance ~ prepostmatch, data = uuf.kid.il, perm=9999) #p=0.1513
uuf.kiddog.ch<-filter(uuf.kiddog.c, contact=="High") 
uuf.kiddog.cl<-filter(uuf.kiddog.c, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.ch, perm=9999) #p<0.0001
adonis(Distance ~ prepostmatch, data = uuf.kiddog.cl, perm=9999) #p=0.0861
uuf.kiddog.ih<-filter(uuf.kiddog.i, contact=="High") 
uuf.kiddog.il<-filter(uuf.kiddog.i, contact=="Low") 
adonis(Distance ~ prepostmatch, data = uuf.kiddog.ih, perm=9999) #p<0.0001
adonis(Distance ~ prepostmatch, data = uuf.kiddog.il, perm=9999) #p=0.0371


#### WUF ####

####  Get one-way adonis results each analysis - kid-kid
adonis(Distance ~ prepostmatch, data = wuf.kid.full, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kid.full, perm=9999)  
adonis(Distance ~ contact, data = wuf.kid.full, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = wuf.kid.full, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kid.full, perm=9999)  
#stratified - 
## by visittype
wuf.kid.c<-filter(wuf.kid.full, visittype=="Control") 
adonis(Distance ~ prepostmatch, data = wuf.kid.c, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kid.c, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = wuf.kid.c, perm=9999) 
adonis(Distance ~ contact, data = wuf.kid.c, perm=9999) 
wuf.kid.i<-filter(wuf.kid.full, visittype=="Intervention") 
adonis(Distance ~ prepostmatch, data = wuf.kid.i, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kid.i, perm=9999)
adonis(Distance ~ visitprepostmatch, data = wuf.kid.i, perm=9999) 
adonis(Distance ~ contact, data = wuf.kid.i, perm=9999) 
## by contact
wuf.kid.h<-filter(wuf.kid.full, contact=="High") 
adonis(Distance ~ prepostmatch, data = wuf.kid.h, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kid.h, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = wuf.kid.h, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kid.h, perm=9999) 
wuf.kid.l<-filter(wuf.kid.full, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kid.l, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kid.l, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = wuf.kid.l, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kid.l, perm=9999) 

####  Get one-way adonis results each analysis - kid-dog
adonis(Distance ~ prepostmatch, data = wuf.kiddog.full, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kiddog.full, perm=9999)  
adonis(Distance ~ contact, data = wuf.kiddog.full, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = wuf.kiddog.full, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kiddog.full, perm=9999)  
#stratified - 
## by visittype
wuf.kiddog.c<-filter(wuf.kiddog.full, visittype=="Control") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.c, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kiddog.c, perm=9999) 
adonis(Distance ~ visitprepostmatch, data = wuf.kiddog.c, perm=9999) 
adonis(Distance ~ contact, data = wuf.kiddog.c, perm=9999) 
wuf.kiddog.i<-filter(wuf.kiddog.full, visittype=="Intervention") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.i, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kiddog.i, perm=9999)
adonis(Distance ~ visitprepostmatch, data = wuf.kiddog.i, perm=9999) 
adonis(Distance ~ contact, data = wuf.kiddog.i, perm=9999) 
## by contact
wuf.kiddog.h<-filter(wuf.kiddog.full, contact=="High") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.h, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kiddog.h, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = wuf.kiddog.h, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kiddog.h, perm=9999) 
wuf.kiddog.l<-filter(wuf.kiddog.full, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.l, perm=9999) 
adonis(Distance ~ visitmatch, data = wuf.kiddog.l, perm=9999)  
adonis(Distance ~ visitprepostmatch, data = wuf.kiddog.l, perm=9999) 
adonis(Distance ~ visittype, data = wuf.kiddog.l, perm=9999) 

### multivariate adonis kid-kid
adonis(Distance ~ prepostmatch + visitmatch, data=wuf.kid.full, perm=9999) 
adonis(Distance ~ prepostmatch * visitmatch, data=wuf.kid.full, perm=9999) 
adonis(Distance ~ prepostmatch + visitmatch + contact + visittype, data=wuf.kid.full, perm=9999)
### multivariate adonis kid-dog
adonis(Distance ~ prepostmatch + visitmatch, data=wuf.kiddog.full, perm=9999) 
adonis(Distance ~ prepostmatch * visitmatch, data=wuf.kiddog.full, perm=9999)
adonis(Distance ~ prepostmatch + visitmatch + contact + visittype, data=wuf.kiddog.full, perm=9999)

## Double Stratified Adonis 
wuf.kid.ch<-filter(wuf.kid.c, contact=="High") 
wuf.kid.cl<-filter(wuf.kid.c, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kid.ch, perm=9999)  #p=0.0005
adonis(Distance ~ prepostmatch, data = wuf.kid.cl, perm=9999)  #p=0.0128
wuf.kid.ih<-filter(wuf.kid.i, contact=="High") 
wuf.kid.il<-filter(wuf.kid.i, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kid.ih, perm=9999) #p=0.3408
adonis(Distance ~ prepostmatch, data = wuf.kid.il, perm=9999) #p=0.4471
wuf.kiddog.ch<-filter(wuf.kiddog.c, contact=="High") 
wuf.kiddog.cl<-filter(wuf.kiddog.c, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.ch, perm=9999) #p=0.0919
adonis(Distance ~ prepostmatch, data = wuf.kiddog.cl, perm=9999) #p=0.1717
wuf.kiddog.ih<-filter(wuf.kiddog.i, contact=="High") 
wuf.kiddog.il<-filter(wuf.kiddog.i, contact=="Low") 
adonis(Distance ~ prepostmatch, data = wuf.kiddog.ih, perm=9999) #p<0.0001
adonis(Distance ~ prepostmatch, data = wuf.kiddog.il, perm=9999) #p=0.0005


#########################################################################################################

### Get Change Variable Post - Pre Distance
# need to match on both subject1 and subject2 (in case of kiddog uniqueid1 & 2)


uuf.kid.pre<-filter(uuf.kid.full, prepostmatch=="PRE") #703x44
uuf.kid.post<-filter(uuf.kid.full, prepostmatch=="POST") #820x44
mergepp<- merge(uuf.kid.pre, uuf.kid.post, by = "subjectid.1", suffixes=c(".pre", ".post")) #18912x87
uuf.kid.chg<- filter(mergepp, subjectid.2.pre==subjectid.2.post) #703x87
uuf.kid.chg$distchange<-uuf.kid.chg$Distance.post-uuf.kid.chg$Distance.pre
uuf.kid.chg1<-filter(uuf.kid.chg, !visittype.pre=="mismatch")
summary(uuf.kid.chg$distchange)
ggplot(uuf.kid.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(aes(x=contact.pre), width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.015, stackdir = "center") +
  facet_grid(.~visittype.pre) +
  ggtitle("UUF Kid-Kid")
uuf.kiddog.pre<-filter(uuf.kiddog.full, prepostmatch=="PRE") #703x44
uuf.kiddog.post<-filter(uuf.kiddog.full, prepostmatch=="POST") #820x44
mergepp2<- merge(uuf.kiddog.pre, uuf.kiddog.post, by = "uniqueid.1", suffixes=c(".pre", ".post")) #18912x87
uuf.kiddog.chg<- filter(mergepp2, uniqueid.2.pre==uniqueid.2.post) #703x87
uuf.kiddog.chg$distchange<-uuf.kiddog.chg$Distance.post-uuf.kiddog.chg$Distance.pre
uuf.kiddog.chg1<-filter(uuf.kiddog.chg, !visittype.pre=="mismatch")
summary(uuf.kiddog.chg$distchange)
ggplot(uuf.kiddog.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.015, stackdir = "center") +
  facet_grid(.~visittype.pre)+
  ggtitle("UUF Kid-Dog")

wuf.kid.pre<-filter(wuf.kid.full, prepostmatch=="PRE") #703x44
wuf.kid.post<-filter(wuf.kid.full, prepostmatch=="POST") #820x44
mergepp<- merge(wuf.kid.pre, wuf.kid.post, by = "subjectid.1", suffixes=c(".pre", ".post")) #18912x87
wuf.kid.chg<- filter(mergepp, subjectid.2.pre==subjectid.2.post) #703x87
wuf.kid.chg$distchange<-wuf.kid.chg$Distance.post-wuf.kid.chg$Distance.pre
wuf.kid.chg1<-filter(wuf.kid.chg, !visittype.pre=="mismatch")
summary(wuf.kid.chg$distchange)
ggplot(wuf.kid.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(aes(x=contact.pre), width=0.95) +
  facet_grid(.~visittype.pre)+
  geom_dotplot(binaxis = "y", binwidth = 0.03, stackdir = "center") +
  ggtitle("WUF Kid-Kid")
wuf.kiddog.pre<-filter(wuf.kiddog.full, prepostmatch=="PRE") #703x44
wuf.kiddog.post<-filter(wuf.kiddog.full, prepostmatch=="POST") #820x44
mergepp2<- merge(wuf.kiddog.pre, wuf.kiddog.post, by = "uniqueid.1", suffixes=c(".pre", ".post")) #18912x87
wuf.kiddog.chg<- filter(mergepp2, uniqueid.2.pre==uniqueid.2.post) #703x87
wuf.kiddog.chg$distchange<-wuf.kiddog.chg$Distance.post-wuf.kiddog.chg$Distance.pre
wuf.kiddog.chg1<-filter(wuf.kiddog.chg, !visittype.pre=="mismatch")
summary(wuf.kiddog.chg$distchange)
ggplot(wuf.kiddog.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.03, stackdir = "center") +
  facet_grid(.~visittype.pre)+
  ggtitle("WUF Kid-Dog")



#####################################################################################################################
### PLOTTING BETA DISTANCE POST-PRE CHANGE

## usign these as main graph, then all samples as supplement

mytheme<-theme_bw() +
  theme(panel.background = element_rect(colour = "white"), 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size=12), 
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=12), 
        strip.background = element_rect(fill="light blue"), 
        strip.text = element_text(size=12, face="bold"),
        plot.title = element_text(size=14, face="bold", hjust=0.5)
        )

uuf.kk<-ggplot(uuf.kid.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(aes(x=contact.pre), width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.02, stackdir = "center") +
  facet_grid(.~visittype.pre) +
  scale_fill_manual(values = colors_contact) +
  geom_hline(yintercept = 0, size=0.8, color="black", linetype="dashed") +
  guides(fill=FALSE) +
  ggtitle("A. Unweighted UniFrac Kid-Kid Distance") +
  scale_y_continuous(name= "Beta Distance", 
                     limits=c(-0.5, 0.55), 
                     breaks=c(-0.45, -0.25, 0, 0.25, 0.5), 
                     labels=c("More\nSimilar", "", "", "", "Less\nSimilar")) +
  mytheme

wuf.kid.chg12<-filter(wuf.kid.chg1, distchange>=-0.5)
wuf.kk<-ggplot(wuf.kid.chg12, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(aes(x=contact.pre), width=0.95) +
  facet_grid(.~visittype.pre) +
  geom_dotplot(binaxis = "y", binwidth = 0.02, stackdir = "center") +
  guides(fill=FALSE) +
  scale_fill_manual(values = colors_contact) +
  geom_hline(yintercept = 0, size=0.8, color="black", linetype="dashed") +
  ggtitle("B. Weighted UniFrac Kid-Kid Distance") + 
  scale_y_continuous(name= "Beta Distance", 
                     limits=c(-0.5, 0.55), 
                     breaks=c(-0.45, -0.25, 0, 0.25, 0.5), 
                     labels=c("More\nSimilar", "", "", "", "Less\nSimilar")) +
  mytheme 

uuf.kd<-ggplot(uuf.kiddog.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.025, stackdir = "center") +
  facet_grid(.~visittype.pre)+
  guides(fill=FALSE) +
  labs(y="Beta Distance") +
  scale_fill_manual(values = colors_contact) +
  geom_hline(yintercept = 0, size=0.8, color="black", linetype="dashed") +
  ggtitle("C. Unweighted UniFrac Kid-Dog Distance") +
  scale_y_continuous(name= "Beta Distance", 
                     limits=c(-0.6, 0.6), 
                     breaks=c(-0.5, -0.25, 0, 0.25, 0.5), 
                     labels=c("More\nSimilar", "", "", "", "Less\nSimilar")) +
  mytheme

wuf.kd<-ggplot(wuf.kiddog.chg1, aes(y=distchange, x=contact.pre, fill=contact.pre)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.025, stackdir = "center") +
  facet_grid(.~visittype.pre)+
  guides(fill=FALSE) +
  labs(y="Beta Distance") +
  scale_fill_manual(values = colors_contact) +
  geom_hline(yintercept = 0, size=0.8, color="black", linetype="dashed") +
  ggtitle("D. Weighted UniFrac Kid-Dog Distance") +
  scale_y_continuous(name= "Beta Distance", 
                     limits=c(-0.6, 0.6), 
                     breaks=c(-0.5, -0.25, 0, 0.25, 0.5), 
                     labels=c("More\nSimilar", "", "", "", "Less\nSimilar")) +
  mytheme


### add p values
ukk_ch <- data.frame(contact.pre = "High", distchange=0.51, 
                         lab = "p=0.0001",
                     visittype.pre = factor("Control", 
                                               levels = c("Control","Intervention")))
ukk_cl <- data.frame(contact.pre = "Low", distchange=0.51, 
                       lab = "p=0.6924",
                     visittype.pre = factor("Control", 
                                             levels = c("Control","Intervention")))
ukk_ih <- data.frame(contact.pre = "High", distchange=0.51, 
                       lab = "p=0.0003",
                     visittype.pre = factor("Intervention", 
                                             levels = c("Control","Intervention")))
ukk_il <- data.frame(contact.pre = "Low", distchange=0.51, 
                       lab = "p=0.1513",
                     visittype.pre = factor("Intervention", 
                                             levels = c("Control","Intervention")))
uuf.kk<-uuf.kk + 
  geom_text(data = ukk_ch, label="p=0.0001", size=3, fontface="bold", color="darkred") +
  geom_text(data = ukk_cl, label="p=0.6924", size=3, color="black") +
  geom_text(data = ukk_ih, label="p=0.0003", size=3, fontface="bold", color="darkred") +
  geom_text(data = ukk_il, label="p=0.1513", size=3, color="black")
wkk_ch <- data.frame(contact.pre = "High", distchange=0.51, 
                     lab = "p=0.0005",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
wkk_cl <- data.frame(contact.pre = "Low", distchange=0.51, 
                     lab = "p=0.0128",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
wkk_ih <- data.frame(contact.pre = "High", distchange=0.51, 
                     lab = "p=0.3408",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
wkk_il <- data.frame(contact.pre = "Low", distchange=0.51, 
                     lab = "p=0.3408",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
wuf.kk<-wuf.kk + 
  geom_text(data = wkk_ch, label="p=0.0005", size=3, fontface="bold", color="darkred") +
  geom_text(data = wkk_cl, label="p=0.0128", size=3, color="black") +
  geom_text(data = wkk_ih, label="p=0.3408", size=3, color="black") +
  geom_text(data = wkk_il, label="p=0.3408", size=3, color="black")
ukd_ch <- data.frame(contact.pre = "High", distchange=0.51, 
                     lab = "p=0.0001",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
ukd_cl <- data.frame(contact.pre = "Low", distchange=0.51, 
                     lab = "p=0.0861",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
ukd_ih <- data.frame(contact.pre = "High", distchange=0.51, 
                     lab = "p=0.0001",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
ukd_il <- data.frame(contact.pre = "Low", distchange=0.51, 
                     lab = "p=0.0371",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
uuf.kd<-uuf.kd + 
  geom_text(data = ukd_ch, label="p=0.0001", size=3, fontface="bold", color="darkred") +
  geom_text(data = ukd_cl, label="p=0.0861", size=3, color="black") +
  geom_text(data = ukd_ih, label="p=0.0001", size=3, fontface="bold", color="darkred") +
  geom_text(data = ukd_il, label="p=0.0371", size=3, color="black")
wkd_ch <- data.frame(contact.pre = "High", distchange=0.60, 
                     lab = "p=0.0919",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
wkd_cl <- data.frame(contact.pre = "Low", distchange=0.60, 
                     lab = "p=0.1717",
                     visittype.pre = factor("Control", 
                                            levels = c("Control","Intervention")))
wkd_ih <- data.frame(contact.pre = "High", distchange=0.51, 
                     lab = "p=0.0001",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
wkd_il <- data.frame(contact.pre = "Low", distchange=0.51, 
                     lab = "p=0.0005",
                     visittype.pre = factor("Intervention", 
                                            levels = c("Control","Intervention")))
wuf.kd<-wuf.kd + 
  geom_text(data = wkd_ch, label="p=0.0919", size=3, color="black") +
  geom_text(data = wkd_cl, label="p=0.1717", size=3, color="black") +
  geom_text(data = wkd_ih, label="p=0.0001", size=3, fontface="bold", color="darkred") +
  geom_text(data = wkd_il, label="p=0.0005", size=3, fontface="bold", color="darkred")




ggarrange(uuf.kk, wuf.kk, uuf.kd, wuf.kd, ncol=2, nrow=2, legend="none")




#################################################################################################################

###### PLOTTING BETA DISTANCES - All ##########

## distance files: wuf.kid.full, wuf.kiddog.full, wuf.kid.full, wuf.kiddog.full
## goal is to plot distance by pre post and make 3 separate graphs (all samples, stratified by visittype, stratified by contact)
## then combine for one big graph
## do for wuf + WUF kid-kid, then wuf + WUF kid-dog


##### Unweighted ######
#### KID - KID Distance ####
## all samples
colors.all<-c("lightblue")
kkall<-ggplot(uuf.kid.full, aes(y=Distance, x=prepostmatch, fill=study.1)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.0075, stackdir = "center") +
  ylab("Unweighted Unifrac Distance") +
  scale_fill_manual(values = colors.all) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                   labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )
kkall
## by visit type - remove mistmatch visits since don't care about them
uuf.kid.vmatch<-filter(uuf.kid.full, visittype=="Control" | visittype=="Intervention")
colors.visit<-c("#1f78b4", "#fc8d62")
kkvisit<-ggplot(uuf.kid.vmatch, aes(y=Distance, x=prepostmatch, fill=visittype)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.01, stackdir = "center", position = position_dodge(0.8)) +
  ylab("Unweighted Unifrac Distance") +
  scale_fill_manual(values = colors.visit) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )
# by contact
colors.contact<-c("#7fc97f", "#bc80bd")
kkcontact<-ggplot(uuf.kid.full, aes(y=Distance, x=prepostmatch, fill=contact)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.0075, stackdir = "center", position = position_dodge(0.8))+
  ylab("Unweighted Unifrac Distance") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.contact) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )



### KID - DOG Distance ###
## all samples
colors.all<-c("lightblue")
kdall<-ggplot(uuf.kiddog.full, aes(y=Distance, x=prepostmatch, fill=study.1)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.0075, stackdir = "center") +
  ylab("Unweighted Unifrac Distance") +
  labs(fills="All Samples") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  guides(fill=guide_legend(title = NULL)) +
  scale_fill_manual(values = colors.all, labels=c("All Samples")) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"), 
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )
## by visit type - remove mistmatch visits since don't care about them
uuf.kiddog.vmatch<-filter(uuf.kiddog.full, visittype=="Control" | visittype=="Intervention")
kdvisit<-ggplot(uuf.kiddog.vmatch, aes(y=Distance, x=prepostmatch, fill=visittype)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.01, stackdir = "center", position = position_dodge(0.8)) +
  ylab("Unweighted Unifrac Distance") +
  labs(fill="By Visit Type") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.visit) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"), 
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )
# by contact
kdcontact<-ggplot(uuf.kiddog.full, aes(y=Distance, x=prepostmatch, fill=contact)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.0075, stackdir = "center", position = position_dodge(0.8))+
  ylab("Unweighted Unifrac Distance") +
  labs(fill="By Contact Level") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.contact) +
  scale_y_continuous(breaks=c(0.35,  0.6, 0.9), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )

## Arrange plots together ###
kid.kid.label<-paste("Kid - Kid", "Distance", sep=" ")
kid.dog.label<-paste("Kid - Dog", "Distance", sep=" ")

kid.kid.plots<-ggarrange(kkall, kkvisit+rremove("y.text")+rremove("y.title"), kkcontact+rremove("y.text")+rremove("y.title"), ncol=3, legend="none")
kid.dog.plots<-ggarrange(kdall, kdvisit+rremove("y.text")+rremove("y.title"), kdcontact+rremove("y.text")+rremove("y.title"), ncol=3, legend="top")

ggdraw() +
  draw_text(kid.kid.label, x = 0.5, y = 0.98, size=16, fontface="bold") +
  draw_plot(kid.kid.plots, x = 0, y = 0.53, width = 1, height = 0.43) +
  draw_text(kid.dog.label, x = 0.5, y = 0.52, size=16, fontface="bold") +
  draw_plot(kid.dog.plots, x = 0, y = 0, width = 1, height = 0.5)

### Alterative without text (better for presentations)
ggdraw() +
  draw_plot(kid.kid.plots, x = 0, y = 0.53, width = 1, height = 0.47) +
  draw_plot(kid.dog.plots, x = 0, y = 0, width = 1, height = 0.52)


#################################################################################################

## Distance for Weighted


#### KID - KID Distance ####
## all samples

kkall.w<-ggplot(wuf.kid.full, aes(y=Distance, x=prepostmatch, fill=study.1)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.016, stackdir = "center") +
  ylab("Weighted UniFrac Distance") +
  scale_fill_manual(values = colors.all) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )
kkall.w
## by visit type - remove mistmatch visits since don't care about them
wuf.kid.vmatch<-filter(wuf.kid.full, visittype=="Control" | visittype=="Intervention")

kkvisit.w<-ggplot(wuf.kid.vmatch, aes(y=Distance, x=prepostmatch, fill=visittype)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.022, stackdir = "center", position = position_dodge(0.8)) +
  ylab("Weighted Unifrac Distance") +
  scale_fill_manual(values = colors.visit) +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4),  
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )
# by contact
kkcontact.w<-ggplot(wuf.kid.full, aes(y=Distance, x=prepostmatch, fill=contact)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.014, stackdir = "center", position = position_dodge(0.8))+
  ylab("Weighted Unifrac Distance") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.contact) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4),  
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold")
  )



### KID - DOG Distance ###
## all samples

kdall.w<-ggplot(wuf.kiddog.full, aes(y=Distance, x=prepostmatch, fill=study.1)) +
  geom_boxplot(width=0.95) +
  geom_dotplot(binaxis = "y", binwidth = 0.018, stackdir = "center") +
  ylab("Weighted UniFrac Distance") +
  labs(fills="All Samples") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  guides(fill=guide_legend(title = NULL)) +
  scale_fill_manual(values = colors.all, labels=c("All Samples")) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4),   
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"), 
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )
## by visit type - remove mistmatch visits since don't care about them
wuf.kiddog.vmatch<-filter(wuf.kiddog.full, visittype=="Control" | visittype=="Intervention")
kdvisit.w<-ggplot(wuf.kiddog.vmatch, aes(y=Distance, x=prepostmatch, fill=visittype)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.025, stackdir = "center", position = position_dodge(0.8)) +
  ylab("Weighted Unifrac Distance") +
  labs(fill="By Visit Type") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.visit) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4),
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"), 
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )
# by contact
kdcontact.w<-ggplot(wuf.kiddog.full, aes(y=Distance, x=prepostmatch, fill=contact)) +
  geom_boxplot(width=1) +
  geom_dotplot(binaxis = "y", binwidth = 0.02, stackdir = "center", position = position_dodge(0.8))+
  ylab("Weighted Unifrac Distance") +
  labs(fill="By Contact Level") +
  scale_x_discrete(limits=c("PRE", "POST")) +
  scale_fill_manual(values = colors.contact) +
  scale_y_continuous(limits = c(0, 1.5), 
                     breaks=c(0.2,  0.8, 1.4), 
                     labels=c("More\nSimilar", "", "Less\nSimilar")) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),  
    axis.text.x = element_text(size=11), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    plot.title = element_text(hjust = 0.5, size=16, face="bold"),
    legend.background = element_rect(colour="black"),
    legend.position = "top", 
    legend.title = element_text(size=11)
  )

## Arrange plots together ###
kid.kid.plots.w<-ggarrange(kkall.w, kkvisit.w+rremove("y.text")+rremove("y.title"), kkcontact.w+rremove("y.text")+rremove("y.title"), ncol=3, legend="none")
kid.dog.plots.w<-ggarrange(kdall.w, kdvisit.w+rremove("y.text")+rremove("y.title"), kdcontact.w+rremove("y.text")+rremove("y.title"), ncol=3, legend="top")


ggdraw() +
  draw_text(kid.kid.label, x = 0.5, y = 0.98, size=16, fontface="bold") +
  draw_plot(kid.kid.plots.w, x = 0, y = 0.53, width = 1, height = 0.43) +
  draw_text(kid.dog.label, x = 0.5, y = 0.52, size=16, fontface="bold") +
  draw_plot(kid.dog.plots.w, x = 0, y = 0, width = 1, height = 0.5)

### Alterative without text (better for presentations)
ggdraw() +
  draw_plot(kid.kid.plots.w, x = 0, y = 0.53, width = 1, height = 0.47) +
  draw_plot(kid.dog.plots.w, x = 0, y = 0, width = 1, height = 0.52)



```





